---
title: "Owl Limpet Visualization 2023-2024"
author: "Cindy Fajardo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../Output/")
```

#Welcome
In this file, I am creating visualization graphs of population size in the seasons beginning with Fall 2023. It also size distribution of tagged limpets and graphs depicting growth.

#Libraries
```{r, Libraries}
library("rerddap")
library(tidyverse)
library("tidync")
library("doParallel")
library(lubridate)
library(patchwork)
library(viridis)
library(here)
library(kableExtra)
library(hrbrthemes)
library(emmeans)
library(agricolae)
library(vegan)
library(performance)
library(broom)
library(car)
library(lme4)
library(lmerTest)
library(pscl)
library(ggridges)
library(devtools)
library(pairwiseAdonis)
```

#Read in CSV
In this chunk, I am reading in the csv's that are relevant to growth data ONLY
```{r, Read in CSV}
feb_cnm_shell_length <- read_csv(here("Data","2024_02_CNM_OL_SL_Data.csv"))
master_sl <- read_csv(here("Data","Master_CNM_OL_SL_Data.csv")) %>%
mutate(sex=ifelse(SL_mm>= 50, "female", "male"))
```


#Simple Statistics
In this chunk, I am looking a simple stats like mean, max, min, standard deviation, standard error, variance, etc
```{r, Simple Statistics}
winter_data <- filter(master_sl, Season=="Winter")
feb_mean <- mean(winter_data$SL_mm)
fall_data <- filter(master_sl, Season=="Fall")
september_mean <- mean(fall_data$SL_mm)
max(feb_cnm_shell_length$SL_mm)
min(feb_cnm_shell_length$SL_mm)

master_sl
```

#Population Shell Length Distribution Graph for Winter
In this chunk, I am creating a bar graph showing the shell lenght distribution for the owl limpets found in my plots in zone 3 for data collection in February 2024
```{r, Population Shell Length Distribution Graph}
feb_cnm_pop_density <- feb_cnm_shell_length %>% #calling csv with shell length data
  select(Limpet_No, SL_mm) %>% #selecting limpet number to count and shell length
  ggplot( aes(x=SL_mm)) + #Sl on x, I am not selecting y so R can use limpet count for density plot
  geom_bar(fill="blue",#color of bars
           colour="black",#color of bar outline
           alpha= 1,#transparency of bar
           width = 1) + #size of bars
  labs(x = "Shell Length (mm)")+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_blank())+
  scale_x_continuous(breaks = c(10,20,30,40,50,60,70,80))+ #lines I want to show up on x axis
  scale_y_continuous(limits = c(0,90))+ #lines I want to show up on y axis
  geom_vline(xintercept = mean(feb_cnm_shell_length$SL_mm),
             colour="green")+ #draw line where average lies on density
  theme_bw()
 feb_cnm_pop_density
```

#Size Distribution of Tagged Limpets
In this chunk, I am creating a new data frame with tagged limpets ONLY
```{r, Size Distribution of Tagged Limpets}
feb_tagged_limpets <- feb_cnm_shell_length %>% #named new data frame using shell length data
  filter(Tag_ID != 0) #filtered out non-tagged limpets
feb_tagged_limpets #calling data frame
mean(feb_tagged_limpets$SL_mm)
```

#Tagged Limpets Shell Length Distribution Graph
This chunk gives population size distribution of tagged limpets in winter
```{r, #Tagged Limpets Shell Length Distribution Graph}
feb_tagged_distribution <- feb_tagged_limpets %>% #calling data frame with tagged limpets only
  select(Limpet_No, SL_mm) %>% #selecting limpet number to count and shell length
  ggplot( aes(x=SL_mm)) + #Sl on x, I am not selecting y so R can use limpet count for density plot
  geom_bar(fill="red",#color of bars
           colour="black",#color of bar outline
           alpha= 1,#transparency of bar
           width = 1) + #size of bars
  labs(title = "Size Distribution of Tagged Owl Limpets, Cabrillo National Monument",
       x = "Shell Length (mm)",
       y = "Owl Limpet Count")+
  theme(plot.title = element_text(size = 9),
        axis.title.y = element_text(size = 8),
        axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7))+
  scale_y_continuous(breaks = c(0,2,4,6,8,10))+ #lines I want to show up on y axis
  scale_x_continuous(breaks = c(10,20,30,40,50,60,70,80))+ #lines I want to show up on x axis
  geom_vline(xintercept = mean(feb_tagged_limpets$SL_mm),
             colour="green")#draw line where average lies on density
feb_tagged_distribution
```

#Total number of tagged limpets
```{r, Total number of tagged limpets}
feb_tagged_limpets %>%
  group_by(Plot_Name) %>% 
  count(Tag_ID)
```

#February Size Density Graph
Population size density graph
```{r, February Size Density Graph}
p <- feb_cnm_shell_length %>%
  ggplot( aes(x=SL_mm)) +
    geom_density(fill = "#FF6666", alpha = 0.6)+
  labs(title = "February")+
    theme_ipsum()

p
```

#February Density Graph
Histogram of population size in February
```{r, February Density Graph}
r <- feb_cnm_shell_length %>%
  ggplot( aes(x=SL_mm)) +
    geom_histogram(binwidth = 3, fill = "#70070e", color= "black")+
  labs(title = "February")+
    theme_ipsum()
r
```

#Septemeber Size Density Graph
Density graph of population size in September
```{r, Septemeber size Density Graph}
q <- fall_data %>%
  ggplot( aes(x=SL_mm)) +
    geom_density(fill = "#52B4C9", alpha = 0.6)+
  labs(title = "September")+
    theme_ipsum()

q
```

#Density of September
Histogram of graph of population distribution in september
```{r, Density of September Graph}
s <- fall_data %>%
  ggplot( aes(x=SL_mm)) +
    geom_histogram(binwidth = 3, fill = "#bf630d", color= "black")+
  labs(title = "September")+
    theme_ipsum()
s
```

#Patchwork of September and February Shell Length
Density graphs of both September and February
```{r, Patchwork of September and February Shell Length}
q+p+
  plot_annotation(title = 'Shell Lenght Density of Owl Limpets in September and February', theme = theme(plot.title = element_text(size = 12)))
```

#Patchwork of Fall and Winter Density
Histograms of Fall and winter
```{r, Patwork of Fall and Winter Density}
s+r
```

#Growth Rates
In this portion I will begin to visualize growth between the seasons

#Population Average Growth rate of Tagged Limpets from September to February
```{r, Growth rate of Tagged Limpets from September to February}
by_sex_data <- master_sl %>%
  filter(Tag_ID!="0") %>% 
  group_by(sex, Tag_ID) %>%
  reframe(limpet_growth = (SL_mm[Season=="Winter"]-SL_mm[Season=="Fall"])/4.75) %>% 
  left_join(master_sl %>% 
              filter(Tag_ID!="0",
                     Season=="Fall") %>% 
              select(Tag_ID,SL_mm))

ggplot(by_sex_data, aes(x=SL_mm,
                        y = limpet_growth,
                        color = sex))+
         geom_boxplot()

 Density_Data <- master_sl %>%
   group_by(Season, Plot_Name)%>%
   count()%>%
   mutate(limpet_density_m2 = n/20)
 
 # run model for dependent variable ~ independent, then for normality, etc, interpret model
 mod_density_season <-lm(limpet_density_m2~Season, data = Density_Data)
check_model(mod_density_season)
anova(mod_density_season) 
summary(mod_density_season)

ggplot(Density_Data, aes(x=Plot_Name,
                         y=limpet_density_m2,
                       color=Season))+
         geom_boxplot()+
  scale_color_brewer(palette = "Set1")


 growth_sl <- master_sl %>% 
  filter(Date != "2024_02",
         Tag_ID !=0) %>% 
  unite(season_tag, c(Season, Tag_ID))
growth_sl

initial_new <- (feb_mean - september_mean)
day_growth_rate <- initial_new/72 #rate per day
day_growth_rate
week_growth_rate<- initial_new/19
week_growth_rate
month_grow_rate <- initial_new/4.75
month_grow_rate
```

#Boxplot of Raw Density
Boxplot of raw density owl limpet population in zone three by season
```{r}
sep_feb_master <- filter(master_sl, Date != "2023_12")

raw_density <- sep_feb_master %>%
  ggplot(aes(x = as.factor(Season), y = SL_mm, fill = Season))+
  geom_boxplot()+
  labs(title = "Raw Density",
       fill = "Season",
       x = "Season")+
  theme(plot.title = element_text(size = 8),
        axis.title.x = element_text(size = 7),
        axis.text.x = element_text(size = 6))
raw_density
```

#Population Density in Fall and Winter
Density graph of owl limpet population per season overlapping each other
```{r, Population Density in Fall and Winter}
m <- master_sl %>%
  ggplot( aes(x=SL_mm)) +
    geom_density(aes(fill = Season), alpha = 0.6)+
  labs(title = "Population Density in Fall and Winter",
       x = "Shell Length",
       y = "Density")+
    theme_ipsum()

m
```