---
title: "Owl Limpet Visualization 2023-2024"
author: "Cindy Fajardo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../Output/")
```

#Welcome
In this file, I am creating visualization graphs of population size in the seasons beginning with Fall 2023. It also size distribution of tagged limpets and graphs depicting growth.

#Libraries
```{r, Libraries}
library("rerddap")
library(tidyverse)
library("tidync")
library("doParallel")
library(lubridate)
library(patchwork)
library(viridis)
library(here)
library(kableExtra)
library(hrbrthemes)
library(emmeans)
library(agricolae)
library(vegan)
library(performance)
library(broom)
library(car)
library(lme4)
library(lmerTest)
library(pscl)
library(ggridges)
library(devtools)
library(pairwiseAdonis)
```

#Read in CSV
In this chunk, I am reading in the csv's that are relevant to growth data ONLY
```{r, Read in CSV}
#feb_cnm_shell_length <- read_csv(here("Data","2024_02_CNM_OL_SL_Data.csv"))
master_sl <- read_csv(here("Data","Master_CNM_OL_SL_Data.csv")) %>%
mutate(sex=ifelse(SL_mm>= 50, "female", "male"))
```


#Filtering Data
Filtered data to respective month of collection
```{r, Simple Statistics}
Winter_data <- filter(master_sl, Season == "Winter")
Fall_data <- master_sl %>% 
  select(Season, Date, Plot_Name, SL_mm, Tag_ID, sex) %>% 
  filter(Season=="Fall")
Spring_data <- filter(master_sl, Season == "Late_Spring")
March_data <- filter(master_sl, Date =="2024_03")
#June_data <- filter(master_sl, Date == "2024_06")
```

#Simple Statistics
In this chunk, I am looking a simple stats like mean, max, min, standard deviation, standard error, variance, etc
```{r}
Winter_SL_mean <- mean(Winter_data$SL_mm)
Fall_SL_mean <- mean(Fall_data$SL_mm)
Spring_SL_mean <- mean(Spring_data$SL_mm)
#march_SL_mean <- mean(March_data$SL_mm)
#june_SL_mean <- mean(June_data$SL_mm)

max(Fall_data$SL_mm)
min(Fall_data$SL_mm)

max(Winter_data$SL_mm)
min(Winter_data$SL_mm)

max(Spring_data$SL_mm)
min(Spring_data$SL_mm)

#max(March_data$SL_mm)
#min(March_data$SL_mm)

#max(June_data$SL_mm)
#min(June_data$SL_mm)

```


#Population Shell Length Distribution Graph for Winter
In this chunk, I am creating a bar graph showing the shell length distribution for the owl limpets found in my plots in zone 3 for data collection in February 2024
```{r, Population Shell Length Distribution Graph}
Winter_pop_density <- Winter_data %>%
  filter(Date == "2024_02") %>% #calling df with shell length data 
  ggplot( aes(x=SL_mm)) + #Sl on x, I am not selecting y so R can use limpet count for density plot
  geom_bar(fill="blue",#color of bars
           colour="black",#color of bar outline
           alpha= 1,#transparency of bar
           width = 1) + #size of bars
  labs(title = "Shell Length Distribution of Owl Limpets At Cabrillo National Monument in Winter 2024",
         x = "Shell Length (mm)",
       y = "Owl Limpet Count")+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_blank())+
  scale_x_continuous(breaks = c(10,20,30,40,50,60,70,80))+ #lines I want to show up on x axis
  scale_y_continuous(limits = c(0,90))+ #lines I want to show up on y axis
  geom_vline(xintercept = median(Winter_data$SL_mm),
             colour="green")+ #draw line where average lies on density
  theme_bw()
 Winter_pop_density
```

#Population Shell Length Distribution Graph for Fall
```{r, Population Shell Length Distribution Graph for Fall}
sep_cnm_pop_density <- Fall_data %>% #calling df with shell length data 
  ggplot( aes(x=SL_mm)) + #Sl on x, I am not selecting y so R can use limpet count for density plot
  geom_histogram(fill="purple",#color of bars
           colour="black",#color of bar outline
           alpha= 1,#transparency of bar
           binwidth = 1) + #size of bars
  labs(title = "Shell Length Distribution of Owl Limpets At Cabrillo National Monument in Fall 2023",
       x = "Shell Length (mm)",
       y = "Owl Limpet Count")+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_blank())+
  scale_x_continuous(breaks = c(10,20,30,40,50,60,70,80))+ #lines I want to show up on x axis
  scale_y_continuous(limits = c(0,90))+ #lines I want to show up on y axis
  geom_vline(xintercept = median(Fall_data$SL_mm),
             colour="green")+ #draw line where average lies on density
  theme_bw()
 sep_cnm_pop_density
```

#Population Shell Length Distribution Graph for Spring
```{r, Population Shell Length Distribution Graph for Spring}
spring_pop_density <- Spring_data %>% #calling df with shell length data
  select(Limpet_No, SL_mm) %>% #selecting limpet number to count and shell length
  ggplot( aes(x=SL_mm)) + #Sl on x, I am not selecting y so R can use limpet count for density plot
  geom_histogram(fill="pink",#color of bars
           colour="black",#color of bar outline
           alpha= 1,#transparency of bar
           binwidth = 1) + #size of bars
  labs(title = "Shell Length Distribution of Owl Limpets At Cabrillo National Monument in Spring 2024",
       x = "Shell Length (mm)",
       y = "Owl Limpet Count")+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_blank())+
  scale_x_continuous(breaks = c(10,20,30,40,50,60,70,80,90))+
  geom_vline(xintercept = median(Spring_data$SL_mm),
             colour="green")+ #draw line where average lies on density
  theme_bw()
 spring_pop_density
```


#Size Distribution of Tagged Limpets In Winter
In this chunk, I am creating a new data frame with tagged limpets ONLY
```{r, Size Distribution of Tagged Limpets in Winter}
Winter_tagged_limpets <- Winter_data %>% #named new data frame using shell length data
  filter(Date != "2023_12",
         Tag_ID != 0) #filtered out non-tagged limpets
Winter_tagged_limpets #calling data frame
mean(Winter_tagged_limpets$SL_mm)
```

#Analysis of Size Distribution Through The Seasons 
```{r, Analysis of Size Distribution Through The Seasons}
#Data Manipulation
size_dist_all_seasons <-master_sl %>% 
  filter(Season!="Spring",
         Date!="2023_12",
         Plot_Name!="CNM_8")

all_seasons_dist_boxplot <- ggplot(size_dist_all_seasons,
                                   aes(x=Season,
                                       y=SL_mm,
                                       color=Season,
                                       fill=Season))+
  geom_boxplot(show.legend = FALSE)+
  labs(x = "Season",
       y ="Shell Length")+
  scale_y_continuous(breaks = c(20,30,40,50,60,70,80,90))+
  scale_x_discrete(labels=c("Fall","Spring","Winter"))+
  scale_color_brewer(palette = "Set2")+
  scale_fill_brewer(palette = "Set3")+
  theme_bw()
all_seasons_dist_boxplot

#models
SL_dis_mod <- kruskal.test(SL_mm~Season, data=size_dist_all_seasons)
print(SL_dis_mod)
pairwise.wilcox.test(size_dist_all_seasons$SL_mm, size_dist_all_seasons$Season,
                 p.adjust.method = "BH")

```

#Tagged Limpets Shell Length Distribution Graph in Winter
This chunk gives population size distribution of limpets in Winter
```{r, #Tagged Limpets Shell Length Distribution Graph In Winter}
winter_tagged_distribution <- Winter_tagged_limpets %>% #calling data frame with tagged limpets only
  select(Limpet_No, SL_mm) %>% #selecting limpet number to count and shell length
  ggplot( aes(x=SL_mm)) + #Sl on x, I am not selecting y so R can use limpet count for density plot
  geom_bar(fill="red",#color of bars
           colour="black",#color of bar outline
           alpha= 1,#transparency of bar
           width = 1) + #size of bars
  labs(title = "Size Distribution of Tagged Owl Limpets, Cabrillo National Monument",
       x = "Shell Length (mm)",
       y = "Owl Limpet Count")+
  theme(plot.title = element_text(size = 9),
        axis.title.y = element_text(size = 8),
        axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7))+
  scale_y_continuous(breaks = c(0,2,4,6,8,10,12))+ #lines I want to show up on y axis
  scale_x_continuous(breaks = c(20,40,60,80))+ #lines I want to show up on x axis
  geom_vline(xintercept = mean(Winter_tagged_limpets$SL_mm),
             colour="green")#draw line where average lies on density
winter_tagged_distribution
```

#Total number of tagged limpets in Winter
```{r, Total number of tagged limpets In Winter}
Winter_tagged_limpets %>%
  group_by(Plot_Name) %>% 
  count(Tag_ID)
```

#Size Distribution of Tagged Limpets In Fall
```{r, Size Distribution of Tagged Limpets In Fall}
Fall_tagged_limpets <- Fall_data %>% #named new data frame using shell length data
  filter(Tag_ID != 0) #filtered out non-tagged limpets
Fall_tagged_limpets #calling data frame
mean(Fall_tagged_limpets$SL_mm)
```

#Tagged Limpets Shell Length Distribution Graph in Fall
This chunk gives population size distribution of limpets in September
```{r, #Tagged Limpets Shell Length Distribution Graph In Fall}
fall_tagged_distribution <- Fall_tagged_limpets %>% #calling data frame with tagged limpets only 
  ggplot( aes(x=SL_mm)) + #Sl on x, I am not selecting y so R can use limpet count for density plot
  geom_bar(fill="brown",#color of bars
           colour="black",#color of bar outline
           alpha= 1,#transparency of bar
           width = 1) + #size of bars
  labs(title = "Size Distribution of Tagged Owl Limpets, Cabrillo National Monument",
       x = "Shell Length (mm)",
       y = "Owl Limpet Count")+
  theme(plot.title = element_text(size = 9),
        axis.title.y = element_text(size = 8),
        axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7))+
  scale_y_continuous(breaks = c(0,2,4,6,8,10,12,14,16))+ #lines I want to show up on y axis
  scale_x_continuous(breaks = c(20,40,60,80))+ #lines I want to show up on x axis
  geom_vline(xintercept = mean(Fall_tagged_limpets$SL_mm),
             colour="green")#draw line where average lies on density
fall_tagged_distribution
```

#Size Distribution of Tagged Limpets In Spring (June)
```{r, Size Distribution of Tagged Limpets In June}
Spring_tagged_limpets <- Spring_data %>% #named new data frame using shell length data
  filter(Tag_ID != 0) #filtered out non-tagged limpets
Spring_tagged_limpets #calling data frame
mean(Spring_tagged_limpets$SL_mm)
```

#Tagged Limpets Shell Length Distribution Graph in Spring (June)
This chunk gives population size distribution of limpets in Spring
```{r, #Tagged Limpets Shell Length Distribution Graph In Spring}
spring_tagged_distribution <- Spring_tagged_limpets %>% #calling data frame with tagged limpets only 
  ggplot( aes(x=SL_mm)) + #Sl on x, I am not selecting y so R can use limpet count for density plot
  geom_histogram(fill="orange",#color of bars
           colour="black",#color of bar outline
           alpha= 1,#transparency of bar
           binwidth = 1) + #size of bars
  labs(title = "Size Distribution of Tagged Owl Limpets, Cabrillo National Monument",
       x = "Shell Length (mm)",
       y = "Owl Limpet Count")+
  theme(plot.title = element_text(size = 9),
        axis.title.y = element_text(size = 8),
        axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7))+
  scale_y_continuous(breaks = c(0,2,4,6,8,10))+ #lines I want to show up on y axis
  scale_x_continuous(breaks = c(20,40,60,80))+ #lines I want to show up on x axis
  geom_vline(xintercept = mean(Spring_tagged_limpets$SL_mm),
             colour="green")#draw line where average lies on density
spring_tagged_distribution
```

#Size Distribution of Tagged Limpets In early Spring (March)
```{r, Size Distribution of Tagged Limpets In March}
March_tagged_limpets <- master_sl %>% #named new data frame using shell length data
  filter(Date == "2024_03",
         Tag_ID != 0) #filtered out non-tagged limpets
March_tagged_limpets #calling data frame
mean(March_tagged_limpets$SL_mm)
```
#Tagged Limpets Shell Length Distribution Graph in Early Spring (March)
This chunk gives population size distribution of limpets in March which was a re-tagging field trip
```{r, #Tagged Limpets Shell Length Distribution Graph}
march_tagged_distribution <- March_tagged_limpets %>% #calling data frame with tagged limpets only 
  ggplot( aes(x=SL_mm)) + #Sl on x, I am not selecting y so R can use limpet count for density plot
  geom_histogram(fill="green",#color of bars
           colour="black",#color of bar outline
           alpha= 1,#transparency of bar
           binwidth = 1) + #size of bars
  labs(title = "Size Distribution of Tagged Owl Limpets, Cabrillo National Monument",
       x = "Shell Length (mm)",
       y = "Owl Limpet Count")+
  theme(plot.title = element_text(size = 9),
        axis.title.y = element_text(size = 8),
        axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7))+
  scale_y_continuous(breaks = c(0,5,10,15,20))+ #lines I want to show up on y axis
  scale_x_continuous(breaks = c(20,40,60,80))+ #lines I want to show up on x axis
  geom_vline(xintercept = mean(March_tagged_limpets$SL_mm),
             colour="red")#draw line where average lies on density
march_tagged_distribution
```

#Growth Rates
In this portion I will begin to visualize growth between the seasons

#Population Average Growth rate of Tagged Limpets from September to February
```{r, Growth rate of Tagged Limpets from September to February}
by_sex_data <- master_sl %>%
  filter(Tag_ID!="0") %>% 
  group_by(sex, Tag_ID) %>%
  reframe(limpet_growth = (SL_mm[Season=="Winter"]-SL_mm[Season=="Fall"])/4.75) %>% 
  left_join(master_sl %>% 
              filter(Tag_ID!="0",
                     Season=="Fall") %>% 
              select(Tag_ID,SL_mm))

ggplot(by_sex_data, aes(x=SL_mm,
                        y = limpet_growth,
                        color = sex))+
         geom_boxplot()
```

#Density and associated Plots
```{r}
all_seasons_Density_Data <- master_sl %>%
   filter(Season!="Spring", #not including re-tagging event
          Date!="2023_12")  %>% #not including December
   group_by(Season, Plot_Name)%>%
   count()%>%
   mutate(limpet_density_m2 = n/20) %>% 
  mutate(Plot = as.factor(Plot_Name)) %>% 
  mutate(log_density = log(limpet_density_m2),
         Seasons = as.factor(Season))
 
 # run model for dependent variable ~ independent, then for normality, etc, interpret model
mod_density_all_seasons <-lmer(log_density~Seasons+(1|Plot), data = all_seasons_Density_Data)
check_model(mod_density_all_seasons)
anova(mod_density_all_seasons)
summary(mod_density_all_seasons)

all_seasons_density_plot <- ggplot(all_seasons_Density_Data,
                                   aes(x=Season,
                                       y=log_density,
                                       color=Season,
                                       fill=Season))+
  geom_boxplot(show.legend = FALSE)+
  labs(x = "Season",
       y ="Log Limpet Density(#/m²)")+
  scale_x_discrete(labels=c("Fall","Spring","Winter"))+
  scale_color_brewer(palette = "Set2")+
  scale_fill_brewer(palette = "Set3")+
  theme_bw()
  
all_seasons_density_plot  

#Post hoc tests
emmeans(mod_density_all_seasons, list(pairwise~Seasons), adjust = "tukey")


#TukeyHSD(aov(mod_density_all_seasons, adjust = "tukey"))
#(log_density~Season+(1|Plot), data = all_seasons_Density_Data)

###########################
 #growth_sl <- master_sl %>% 
  #filter(Date != "2024_02",
         #Tag_ID !=0) %>% 
  #unite(season_tag, c(Season, Tag_ID))
#growth_sl

#initial_new <- (feb_mean - september_mean)
#day_growth_rate <- initial_new/72 #rate per day
#day_growth_rate
#week_growth_rate<- initial_new/19
#week_growth_rate
#month_grow_rate <- initial_new/4.75
#month_grow_rate
```


#Population Average Growth rate of Tagged Limpets from February to June
This does not include March re-tagging event
```{r, Growth rate of Tagged Limpets from February to June}
Spring_by_sex_data <- master_sl %>%
  filter(Tag_ID!="0",
         Season!= "Spring") %>% 
  group_by(sex, Tag_ID) %>%
  reframe(limpet_growth = (SL_mm[Season=="Winter"]-SL_mm[Season=="Late_Spring"])/4.75) %>% 
  left_join(master_sl %>% 
              filter(Tag_ID!="0",
                     Season=="Late_Spring") %>% 
              select(Tag_ID,SL_mm)) #most values are negative

spring_by_sex_box <- ggplot(Spring_by_sex_data, aes(x=SL_mm,
                        y = limpet_growth,
                        color = sex))+
         geom_boxplot()
spring_by_sex_box

Spring_Density_Data <- master_sl %>%
  filter(Season != "Fall") %>% 
  filter(Season != "Spring",
         Date!="2023_12") %>% 
   group_by(Season, Plot_Name)%>%
   count()%>%
   mutate(limpet_density_m2 = n/50)
   
winter_spring_dens_plot <- ggplot(Spring_Density_Data, 
       aes(x=Plot_Name,
           y=limpet_density_m2,
           color=Season))+
  geom_boxplot()+
  scale_color_brewer(palette = "Set1")
winter_spring_dens_plot

 
 # run model for dependent variable ~ independent, then for normality, etc, interpret model
mod_density_winter_spring <-lm(limpet_density_m2~Season, data = Spring_Density_Data)
check_model(mod_density_winter_spring)
anova(mod_density_winter_spring) 
summary(mod_density_winter_spring)

```

#Population Average Growth rate of Tagged Limpets from March to June
```{r, Growth rate of Tagged Limpets from February to June}
March_by_sex_data <- master_sl %>%
  filter(Tag_ID!="0",
         Season!= "Fall",
         Season!= "Winter") %>% 
  group_by(sex, Tag_ID) %>%
  reframe(limpet_growth = (SL_mm[Season=="Spring"]-SL_mm[Season=="Late_Spring"])/4.75) %>% 
  left_join(master_sl) %>% 
              filter(Tag_ID!="0",
                     Season=="Late_Spring") %>% 
              select(Tag_ID,SL_mm, limpet_growth, sex) %>% 
  filter(limpet_growth > "0")#most values are negative

march_by_sex_box <- ggplot(March_by_sex_data, aes(x=SL_mm,
                        y = limpet_growth,
                        color = sex))+
         geom_boxplot()
march_by_sex_box

March_Density_Data <- master_sl %>%
  filter(Season != "Fall") %>% 
  filter(Season != "Winter") %>% 
   group_by(Season, Plot_Name)%>%
   count()%>%
   mutate(limpet_density_m2 = n/50)
   
March_dens <- ggplot(March_Density_Data, 
       aes(x=Plot_Name,
           y=limpet_density_m2,
           color=Season))+
  geom_boxplot()+
  scale_color_brewer(palette = "Set1")
March_dens

#March was a retagging event so the density will be significantly smaller as we only measured and recorded 50 or less tagged limpets. Cannot compare March to June
```

#February Size Density Graph
Population size density graph
```{r, February Size Density Graph}
p <- Feb_data %>%
  ggplot( aes(x=SL_mm)) +
    geom_density(fill = "#FF6666", alpha = 0.6)+
  labs(title = "February")+
    theme_ipsum()

p
```

#February Density Graph
Histogram of population size in February
```{r, February Density Graph}
r <- Feb_data %>%
  ggplot( aes(x=SL_mm)) +
    geom_histogram(binwidth = 3, fill = "#70070e", color= "black")+
  labs(title = "February")+
    theme_ipsum()
r
```

######Junk under this line, but potentially useful#######
#Septemeber Size Density Graph
Density graph of population size in September
```{r, Septemeber size Density Graph}
q <- Sep_data %>%
  ggplot( aes(x=SL_mm)) +
    geom_density(fill = "#52B4C9", alpha = 0.6)+
  labs(title = "September")+
    theme_ipsum()

q
```

#Density of September
Histogram of graph of population distribution in september
```{r, Density of September Graph}
s <- Sep_data %>%
  ggplot( aes(x=SL_mm)) +
    geom_histogram(binwidth = 3, fill = "#bf630d", color= "black")+
  labs(title = "September")+
    theme_ipsum()
s
```

#Patchwork of September and February Shell Length
Density graphs of both September and February
```{r, Patchwork of September and February Shell Length}
q+p+
  plot_annotation(title = 'Shell Lenght Density of Owl Limpets in September and February', theme = theme(plot.title = element_text(size = 12)))
```

#Patchwork of Fall and Winter Density
Histograms of September and February
```{r, Patwork of Sep and Feb Density}
s+r
```


#Boxplot of Raw Density
Boxplot of raw density owl limpet population in zone three by season
```{r}
sep_feb_master <- filter(master_sl, Date != "2023_12")

raw_density <- sep_feb_master %>%
  ggplot(aes(x = as.factor(Season), y = SL_mm, fill = Season))+
  geom_boxplot()+
  labs(title = "Raw Density",
       fill = "Season",
       x = "Season")+
  theme(plot.title = element_text(size = 8),
        axis.title.x = element_text(size = 7),
        axis.text.x = element_text(size = 6))
raw_density
```

#Population Density in Fall and Winter
Density graph of owl limpet population per season overlapping each other
```{r, Population Density in Fall and Winter}
m <- master_sl %>%
  ggplot( aes(x=SL_mm)) +
    geom_density(aes(fill = Season), alpha = 0.6)+
  labs(title = "Population Density in Fall and Winter",
       x = "Shell Length",
       y = "Density")+
    theme_ipsum()

m
```