---
title: "Cabrillo National Monument: Owl limpet seasonality and community effects"
subtitle: "A mix of univariate of multivariate statistics to test seasonality in both limpet size/density and community responses"
author: "Nyssa/Cindy"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: show
    code-tools: true
    df-print: paged
execute:
  warning: false
  message: false
  cache: false

---

## Overview

This Quarto notebook analyzes three datasets from **Cabrillo National Monument (San Diego, CA)** to answer:

1. **Seasonality in owl limpets (*Lottia gigantea*)**
   - Does **density** change by **Season**?
   - Does **size** (shell length, `SL_mm`) change by **Season**?

2. **Community consequences of limpet presence**
   - Does **owl limpet presence** (`COL_pres`) affect:
     - **Mobile invertebrate composition** (counts)?
     - **Sessile cover composition** (% cover / proportional cover)?
   - Are there **interactions with Season**?

### Design assumptions and why they matter 

- The datasets include repeated sampling across **Transects** and/or **Plots**, so we:
  - use **mixed models** with **random intercepts** for spatial sampling units (e.g., `Transect`, or `Plot_Name`),
  - and constrain permutations by transect for multivariate tests.

- **Density**:
  - The transect size was **20 m²** so I divided the counts by this value to get density per transect .

- **Community data**:
  - Mobile inverts are **counts** (often overdispersed, zero-inflated).
  - Sessile “percent cover” is stored as **proportions** (0–1) in this dataset; we still treat it as compositional.

---

## Packages and helpers

```{r}
# Core
library(tidyverse)
library(here)
library(janitor)     # to clean the names for consistency
library(car)

# Mixed models
library(glmmTMB)     # flexible GLMMs (NB, Gamma, etc.)
library(lme4)        # LMMs (fallback)
library(emmeans)     # estimated marginal means / contrasts

# Diagnostics
library(DHARMa)      # residual simulation diagnostics for GLMMs
library(performance) # assumption checks / R2 / collinearity
library(broom.mixed) # tidy() for mixed models

# Multivariate community ecology
library(vegan)       # adonis2, metaMDS, betadisper, envfit
library(pairwiseAdonis) # optional; pairwise PERMANOVA if needed

# Figures
library(ggdist)      # raincloud/half-eye
library(patchwork)   # figure layouts
library(scales)      # label_percent etc.
library(ggrepel)
library(ggpubr)
library(knitr)
library(kableExtra)
library(webshot2)
library(rerddap)
library(ggmap)


# A consistent plotting theme for a paper
theme_set(
  theme_classic(base_size = 14) + # base size text is 14
    theme(
      panel.grid = element_blank(),
      axis.title = element_text(face = "bold"),
      legend.title = element_text(face = "bold")
    )
)

# Helper: safe logit transform for proportions in [0,1]
# (useful for cover variables with 0s and 1s)
logit01 <- function(p, eps = 1e-4) {
  p2 <- pmin(pmax(p, eps), 1 - eps)
  qlogis(p2)
}
```

---

## Import data

> **Paths** assume the files are in the same folder as this Quarto file, or are in the working directory.
> When you move this notebook to the scripts folder and the data files into your data folder, update paths below. 
> Note, I edited these files to remove the transects removed in your other script and to just bring in the normalized community composition data

```{r}
# File paths (edit if needed)
path_size   <- here("Data", "limpet_size.csv")
path_mobile <- here("Data", "mobile_counts.csv")
path_sess   <- here("Data", "sessile_percent_cover.csv")
path_coordinates <- here("Data", "transect_coordinates.csv")

limpet_size_raw <-  read_csv(path_size, show_col_types = FALSE) %>% clean_names() %>% 
  filter(!season %in% "Spring",
         date != "2023_12_10") %>% 
  mutate(season = ifelse(season == "Late_Spring", "Spring", season),
         size = ifelse(sl_mm >= 30,"large", "small")) # for some reason the season is named Late spring here, but spring everywhere else (This is because when I went out in the winter, I realized that I was loosing tags so we decided to go out and have a re-tagging event in March, or Spring. I re-tagged and measured SL but I didn't do any community surveys, it wasn't part of the plan since I figured it was just about shell length. We then decided to not use the re-tagging event data. I went out again in late May and that is the data that I consider to be the Late_spring data and the only data I use when referring to "Spring". I changed the line of code to reflect that.)
mobile_raw      <- read_csv(path_mobile, show_col_types = FALSE) %>% clean_names()
sessile_raw     <- read_csv(path_sess, show_col_types = FALSE) %>% clean_names()

transect_coordinates <- read_csv(path_coordinates, show_col_types = FALSE) %>% clean_names()

glimpse(limpet_size_raw)
glimpse(mobile_raw)
glimpse(sessile_raw)

# to standardize the colors throughout the manuscript
season_colors <- c( 
  Winter = "#1F4E79",
  Spring = "#2E8B57",
  Fall   = "#C65D2E",
  Summer = "#b30b27")

size_colors <- c(
  large = "#de1899",
  small = "#2a8eeb")

```

---

## Data cleaning and harmonization

We standardize:
- `season` as an **ordered factor**
- `date` as a real date (best-effort parse for differences across datasets)
- `col_pres` to **Yes/No** factor
- key grouping variables (`transect`, `plot_name`, etc.)

```{r}
# Season order 
season_levels <- c("Fall", "Winter","Spring","Summer") # change this order to change order of the plots

# Limpet size dataset
limpet_size <- limpet_size_raw %>%
  mutate(
    season = factor(season, levels = season_levels),
    # Dataset stores date as "YYYY_MM_DD" strings (e.g., 2023_09_29)
    date = parse_date_time(date, orders = c("Y_m_d", "Y_m", "Y-m-d", "Y/m/d")) %>% as_date(),
    plot_name = as.factor(plot_name),
    plot_side = as.factor(plot_side),
    sl_mm = as.numeric(sl_mm),
    size = as.factor(size)
  ) %>%
  filter(!is.na(season), !is.na(sl_mm)) # remove the rows with NAs in length or season

# Mobile invertebrate counts
mobile <- mobile_raw %>%
  mutate(
    season = factor(season, levels = season_levels),
    date = parse_date_time(date, orders = c("Y_m_d", "Y_m", "Y-m-d", "Y/m/d")) %>% as_date(),
    transect = as.factor(transect),
    plot_name = as.factor(plot_name),
    col_pres = case_when(
      tolower(col_pres) %in% c("y","yes","true","t","1") ~ "Yes",
      tolower(col_pres) %in% c("n","no","false","f","0") ~ "No",
      TRUE ~ NA_character_
    ) %>% factor(levels = c("No","Yes"))
  ) %>%
  filter(!transect %in% c("CNM_7", "CNM_8"))

# Sessile percent cover (in this file, cover columns appear to be proportions already: 0–1)
sessile <- sessile_raw %>%
  mutate(
    season = factor(season, levels = season_levels),
    date = parse_date_time(date, orders = c("Y_m_d", "Y_m", "Y-m-d", "Y/m/d")) %>% as_date(),
    transect = as.factor(transect),
    plot_name = as.factor(plot_name),
    col_pres = case_when(
      tolower(col_pres) %in% c("y","yes","true","t","1") ~ "Yes",
      tolower(col_pres) %in% c("n","no","false","f","0") ~ "No",
      TRUE ~ NA_character_
    ) %>% factor(levels = c("No","Yes"))
  )

# Quick checks
limpet_size %>% count(season)
mobile %>% count(season, col_pres) 
sessile %>% count(season, col_pres) ## looks like there are missing pairs of col present
```

---

## Question 1: Do limpet density and size change by season?

### 1A. Density (true density; transect area = 20 m²)

You specified that each limpet-density sampling unit corresponds to a **20 m² transect**.
We therefore compute:

- `n_limpets` = number of limpets observed in a sampling unit
- `density_m2` = `n_limpets / 20`

```{r}
limpet_density <- limpet_size %>%
  group_by(season, date, plot_name, size) %>%
  summarise(
    n_limpets = n(),# count all the limpets
    # True density (per m²) using known transect area
    area_m2 = 20, ## double check this is correct. How long was the transect and how many meters on either size of the transect did you go? 
    density_m2 = n_limpets / area_m2,
    mean_sl = mean(sl_mm, na.rm = TRUE), # mean size per transect
    sd_sl = sd(sl_mm, na.rm = TRUE), # standard deviation in size per transect
    .groups = "drop" # drop the groupings
  )

limpet_density %>% glimpse()

small_density <- limpet_density %>% 
  filter(size == "small")

large_density <- limpet_density %>% 
  filter(size == "large")
  
```

#### Exploratory plot (density by season)

```{r}
p_density <- limpet_density %>%
  ggplot(aes(x = season, y = density_m2, color = season)) +
  ggdist::stat_halfeye(aes(x = season, y = density_m2, fill = season),adjust = 0.7, width = 0.6, justification = -0.15) +
  geom_boxplot(width = 0.18, outlier.shape = NA) + ## this removes the outliers from the boxplot-- only OK because I also have the stat_halfeye and the jitter
  geom_jitter(width = 0.08, height = 0, alpha = 0.35) +
  labs(x = NULL, y = "Owl limpet density (individuals m⁻²)"
       #title = "Seasonal variation in owl limpet density"
       ) +
  scale_color_manual(values = season_colors)+
  scale_fill_manual(values = season_colors)+
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  theme(legend.position = "none")

p_density

#My interpretation: Density is different seasonally. In the fall, density is high all around. There's more plots with 20 or more individuals per meter squared. There's only one plot that has less than 5 individuals per meter squared. That makes sense to me. There were certain plots that the high density was very visible. Plot CNM_6 had the lowest amount of limpets present every season, this was my observation. In the winter, there's a change. The number of individuals per meter squared drops. Three plots have 15+ per meter squared and 3 plots have less than 10. In the spring you can see the addition of recruits to the plots. Spawning occurs in the winter in January/February. Although I am not sure how long the larvae stage is. I can do some research. The Spring data was collected in late spring (Early June), so it gave time for juveniles to find the population and start growing into adults. Again I don't know much about that stage, I'll scower for papers. In these plots, there's a clear addition of individuals per plot consistent with spawning and the abundance of algae due to seasonal upwelling. 
```

#### Exploratory plot (density by season of small individuals)
```{r}
small_density_plot <- small_density %>%
  ggplot(aes(x = season, y = density_m2, color = season)) +
  ggdist::stat_halfeye(aes(x = season, y = density_m2, fill = season),adjust = 0.9, width = 0.6, justification = -0.15) +
  geom_boxplot(width = 0.18, outlier.shape = NA) + ## this removes the outliers from the boxplot-- only OK because I also have the stat_halfeye and the jitter
  geom_jitter(width = 0.08, height = 0, alpha = 0.35) +
  labs(x = NULL, y = "Owl limpet density (individuals m⁻²)"
       #title = "Seasonal variation in owl limpet density"
       ) +
  scale_color_manual(values = season_colors)+
  scale_fill_manual(values = season_colors)+
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.05)))+
  theme(legend.position = "none")

small_density_plot

```

#### Exploratory plot (density by season of large individuals)
```{r}
large_density_plot <- large_density %>%
  ggplot(aes(x = season, y = density_m2, color = season)) +
  ggdist::stat_halfeye(aes(x = season, y = density_m2, fill = season),adjust = 0.7, width = 0.6, justification = -0.15) +
  geom_boxplot(width = 0.18, outlier.shape = NA) + ## this removes the outliers from the boxplot-- only OK because I also have the stat_halfeye and the jitter
  geom_jitter(width = 0.08, height = 0, alpha = 0.35) +
  labs(x = NULL, y = "Owl limpet density (individuals m⁻²)"
       #title = "Seasonal variation in owl limpet density"
       ) +
  scale_color_manual(values = season_colors)+
  scale_fill_manual(values = season_colors)+
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  theme(legend.position = "none")

large_density_plot
```


#### Model: negative binomial GLMM (random effects for plot) for entire population

Counts are typically overdispersed; negative binomial is usually appropriate.

```{r}
m_density <- glmmTMB(
  # Model counts because negative binomial requires count data with an offset for sampled area to estimate density
  n_limpets ~ season + offset(log(area_m2)) + (1 | plot_name), 
  family = nbinom2(),
  data = limpet_density
)

summary(m_density)
performance::check_overdispersion(m_density)
performance::check_zeroinflation(m_density)

```

##### Diagnostics (DHARMa)

```{r}
sim_density <- DHARMa::simulateResiduals(m_density, n = 1000)
plot(sim_density)
DHARMa::testDispersion(sim_density)
DHARMa::testZeroInflation(sim_density)
#https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html
```

##### Effect sizes: estimated marginal means and contrasts

```{r}
emm_density <- emmeans(m_density, ~ season, type = "response")
# With the area offset, 'response' is on the expected count scale; divide by area to report density

emm_density

pairs(emm_density, adjust = "tukey") 
```

##### Figure showing the actual model results: model-based means with uncertainty

```{r}
df_density_emm <- as.data.frame(emm_density) %>%
  mutate(
    area_m2 = 20,
    density = response / area_m2,
    density_lower = asymp.LCL / area_m2,
    density_upper = asymp.UCL / area_m2
  )

# Note add in stars or letters or brackets based on significance testing from the emmeans
p_density_emm <- ggplot(df_density_emm, aes(x = season, y = density)) + 
  ggdist::stat_slab(data = limpet_density, aes(x = season, y = density_m2, fill = season),adjust = 0.7, width = 0.6, justification = -0.15, show.legend = FALSE)+
  geom_jitter(data = limpet_density, aes(x = season, y = density_m2), 
              alpha = 0.5, width = 0.1)+
  geom_point(size = 4) +
  geom_errorbar(data = df_density_emm, aes(ymin = density_lower, ymax = density_upper), width = 0.12) +
  labs(x = NULL, y = "Estimated owl limpet density (individuals/m⁻²)",
       title = "Seasonal effect on owl limpet density (model-based NB-GLMM with area offset)") +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  scale_fill_manual(values = season_colors)
ggsave(filename = "figure3.png", width = 9, height = 6)

p_density_emm

```

#### Model: negative binomial GLMM (random effects for plot) for small individuals

Counts are typically overdispersed; negative binomial is usually appropriate.
```{r}
model_small_density <- glmmTMB(
  # Model counts because negative binomial requires count data with an offset for sampled area to estimate density
  n_limpets ~ season + offset(log(area_m2)) + (1 | plot_name), 
  family = nbinom2(),
  data = small_density
)

summary(model_small_density)
performance::check_overdispersion(model_small_density)
performance::check_zeroinflation(model_small_density)
```

##### Diagnostics (DHARMa) for small individuals

```{r}
small_sim_density <- DHARMa::simulateResiduals(model_small_density, n = 1000)
plot(small_sim_density)
DHARMa::testDispersion(small_sim_density)
DHARMa::testZeroInflation(small_sim_density)
```
##### Effect sizes: estimated marginal means for small individuals

```{r}
small_emm_density <- emmeans(model_small_density, ~ season, type = "response")
# With the area offset, 'response' is on the expected count scale; divide by area to report density

small_emm_density

pairs(small_emm_density, adjust = "tukey")
```

##### Figure showing the actual model results for small individuals: model-based means with uncertainty

```{r}
df_small_density_emm <- as.data.frame(small_emm_density) %>%
  mutate(
    area_m2 = 20,
    density = response / area_m2,
    density_lower = asymp.LCL / area_m2,
    density_upper = asymp.UCL / area_m2
  )

# Note add in stars or letters or brackets based on significance testing from the emmeans
p_small_density_emm <- ggplot(df_small_density_emm, aes(x = season, y = density)) + 
  ggdist::stat_slab(data = small_density, aes(x = season, y = density_m2, fill = season),adjust = 0.7, width = 0.6, justification = -0.15, show.legend = FALSE)+
  geom_jitter(data = small_density, aes(x = season, y = density_m2), 
              alpha = 0.5, width = 0.1)+
  geom_point(size = 4) +
  geom_errorbar(data = df_small_density_emm, aes(ymin = density_lower, ymax = density_upper), width = 0.12) +
  geom_signif(comparisons = list(c("Fall", "Winter")),
              map_signif_level = TRUE, # Automatically uses asterisks for p-values
              annotations = "*")+ 
  geom_signif(comparisons = list(c("Winter", "Spring")),
              map_signif_level = TRUE, # Automatically uses asterisks for p-values
              annotations = "*")+
  labs(x = NULL, y = "Estimated owl limpet density (small individuals/m⁻²)",
       title = "Seasonal effect on small owl limpet density (model-based NB-GLMM with area offset)") +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  scale_fill_manual(values = season_colors)
ggsave(filename = "small.png", width = 9, height = 6)

p_small_density_emm
```

#### Model: negative binomial GLMM (random effects for plot) for large individuals

```{r}
model_large_density <- glmmTMB(
  # Model counts because negative binomial requires count data with an offset for sampled area to estimate density
  n_limpets ~ season + offset(log(area_m2)) + (1 | plot_name), 
  family = nbinom2(),
  data = large_density
)

summary(model_large_density)
performance::check_overdispersion(model_large_density)
performance::check_zeroinflation(model_large_density)
```
##### Diagnostics (DHARMa) for large individuals

```{r}
large_sim_density <- DHARMa::simulateResiduals(model_large_density, n = 1000)
plot(large_sim_density)
DHARMa::testDispersion(large_sim_density)
DHARMa::testZeroInflation(large_sim_density)
```

##### Effect sizes: estimated marginal means for large individuals

```{r}
large_emm_density <- emmeans(model_large_density, ~ season, type = "response")
# With the area offset, 'response' is on the expected count scale; divide by area to report density

large_emm_density

pairs(large_emm_density, adjust = "tukey")
```

##### Figure showing the actual model results for large individuals: model-based means with uncertainty

```{r}
df_large_density_emm <- as.data.frame(large_emm_density) %>%
  mutate(
    area_m2 = 20,
    density = response / area_m2,
    density_lower = asymp.LCL / area_m2,
    density_upper = asymp.UCL / area_m2
  )

# Note add in stars or letters or brackets based on significance testing from the emmeans
p_large_density_emm <- ggplot(df_large_density_emm, aes(x = season, y = density)) + 
  ggdist::stat_slab(data = large_density, aes(x = season, y = density_m2, fill = season),adjust = 0.7, width = 0.6, justification = -0.15, show.legend = FALSE)+
  geom_jitter(data = large_density, aes(x = season, y = density_m2), 
              alpha = 0.5, width = 0.1)+
  geom_point(size = 4) +
  geom_errorbar(data = df_large_density_emm, aes(ymin = density_lower, ymax = density_upper), width = 0.12) +
  geom_signif(comparisons = list(c("Fall", "Winter")),
              map_signif_level = TRUE, # Automatically uses asterisks for p-values
              annotations = "**")+ # Manually set the annotation.
   geom_signif(comparisons = list(c("Fall", "Spring")),
              map_signif_level = TRUE, # Automatically uses asterisks for p-values
              annotations = "***")+
  labs(x = NULL, y = "Estimated owl limpet density (large individuals/m⁻²)",
       title = "Seasonal effect on large owl limpet density (model-based NB-GLMM with area offset)") +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  scale_fill_manual(values = season_colors)
ggsave(filename = "large.png", width = 9, height = 6)

p_large_density_emm
```


```{r}
#| fig-width: 14
#| fig-height: 6
## Bring the figures together for pub quality
## Add in stars with brackes to show significance between seasons based on emmeans
small_large_density <- (p_small_density_emm+labs(title = NULL)|p_large_density_emm+labs(title = NULL))+plot_annotation(tag_levels = "a")
small_large_density
ggsave(filename = "small_large.png", width = 14, height = 6)
#title = "Seasonal effects on Owl Limpet Density and Shell Length"
```
---

### 1B. Size (`sl_mm`) by season

Because `sl_mm` is continuous and strictly positive, we typically choose either:
- **Gamma GLMM** or **Tweedie GLMM** with log link, or
- **LMM** on log-transformed sizes.

Here, we start with **Gamma GLMM** and fallback if needed.
The **Gamma** was not a good fit so I went to *Tweedie* which was better but still not great. Keep looking.... the bimodal sizes makes this harder
What about gaussian with log link or inverse gaussian

```{r}
whole_pop_sl_dist <- limpet_size %>% 
  ggplot(aes(x= sl_mm))+
  geom_histogram(fill= "gray",
                 colour= "black",
                 alpha=1)+
  labs(title = "Shell Length Distribution Sep 2023-June 2024",
         x = "Shell Length (mm)",
       y = "Owl Limpet Count")+
  theme(axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.title.y = element_text(size= 10),
        axis.text.y = element_text(size= 9))+
  geom_vline(xintercept = 34,
             linetype="dashed")
whole_pop_sl_dist
```

```{r}
small_size <- limpet_size %>% 
  filter(size == "small")

small_pop_sl_dist <- small_size %>% 
  ggplot(aes(x= sl_mm))+
  geom_histogram(fill= "gray",
                 colour= "black",
                 alpha=1,
                 binwidth = 2)+
  labs(title = "Shell Length Distribution of Small Individuals Sep 2023-June 2024",
         x = "Shell Length (mm)",
       y = "Owl Limpet Count")+
  theme(axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.title.y = element_text(size= 10),
        axis.text.y = element_text(size= 9))
small_pop_sl_dist
```

```{r}
large_size <- limpet_size %>% 
  filter(size == "large")

large_pop_sl_dist <- large_size %>% 
  ggplot(aes(x= sl_mm))+
  geom_histogram(fill= "gray",
                 colour= "black",
                 alpha=1,
                 binwidth = 2)+
  labs(title = "Shell Length Distribution of Large Individuals Sep 2023-June 2024",
         x = "Shell Length (mm)",
       y = "Owl Limpet Count")+
  theme(axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.title.y = element_text(size= 10),
        axis.text.y = element_text(size= 9))
large_pop_sl_dist
```

```{r}
small_large_size <- limpet_size %>%
  ggplot(aes(x = size, y = sl_mm, color = size)) +
  ggdist::stat_halfeye(adjust = 0.7, width = 0.6, justification = -0.15, aes(fill = size)) +
  geom_boxplot(width = 0.18, outlier.shape = NA) +
  geom_jitter(width = 0.08, height = 0, alpha = 0.10, size = 0.8) +
  scale_fill_manual(values = size_colors)+
  scale_color_manual(values =size_colors)+
  labs(x = NULL, y = "Shell length (mm)",
       title = "Owl limpet size distribution by size"
       ) +
  theme(legend.position = "none")
small_large_size
```

```{r}
leveneTest(sl_mm~size, data = limpet_size)
limpet_size %>%
  group_by(season, size) %>%
  summarise(variance = var(sl_mm),
            mean = mean(sl_mm))

m_size_lmm <- glmmTMB(
  log(sl_mm) ~ size + (1 | plot_name),
  data = limpet_size)

check_model(m_size_lmm)
summary(m_size_lmm)
```

##### Diagnostics

```{r}
small_large_size <- DHARMa::simulateResiduals(m_size_lmm, n = 1000)
plot(small_large_size)
DHARMa::testDispersion(small_large_size)
```

##### Estimated marginal means (back-transformed to mm)

```{r}
emm_small_large <- emmeans(m_size_lmm, ~ size, type = "response")
emm_small_large
pairs(emm_small_large, adjust = "tukey")
```

##### Figure: model-based size means

```{r}
df_small_large_emm <- as.data.frame(emm_small_large)

p_small_large_emm <- ggplot(df_size_emm, aes(x = size, y = response)) +
  ggdist::stat_slab(data = limpet_size, aes(x = size, y = sl_mm, fill = size),adjust = 0.7, width = 0.6, justification = -0.15, alpha = 0.80, show.legend = FALSE)+
  geom_jitter(data = limpet_size, aes(x = size, y = sl_mm, color = size), width = 0.08, height = 0, alpha = 0.10, size = 0.8)+
  geom_errorbar(data = df_small_large_emm, aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.12) +
  scale_fill_manual(values = size_colors)+
  scale_color_manual(values =size_colors)+
  labs(x = NULL, y = "Estimated mean shell length (mm)",
       title = "Size effect on owl limpet size (Gaussian GLMM)"
       ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  theme(legend.position = "none")
#ggsave(filename = "figure4.png", width = 14, height = 6)

p_small_large_emm
```

```{r}
limpet_size %>%
  group_by(season) %>% 
  summarise(n_sl= n())

p_size <- limpet_size %>%
  ggplot(aes(x = season, y = sl_mm, color = season)) +
  ggdist::stat_halfeye(adjust = 0.7, width = 0.6, justification = -0.15, aes(fill = season)) +
  geom_boxplot(width = 0.18, outlier.shape = NA) +
  geom_jitter(width = 0.08, height = 0, alpha = 0.20, size = 0.8) +
  labs(x = NULL, y = "Shell length (mm)",
       title = "Owl limpet size distribution by season"
       ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
   scale_color_manual(values = season_colors)+
  scale_fill_manual(values = season_colors)+
  theme(legend.position = "none")


p_size
```

#### Gamma  or Tweedie GLMM with random effects for plot

```{r}
m_size_tweedie <- glmmTMB(
  sl_mm ~ season + (1 | plot_name),
  dispformula = ~ season,  # Model dispersion as a function of season because unequal variances
  family = tweedie(link = "log"), # better but still not the best, but I trust more this model
  data = limpet_size
)

summary(m_size_tweedie)
```

##### Diagnostics

```{r}
sim_size <- DHARMa::simulateResiduals(m_size_tweedie, n = 1000)
plot(sim_size)
DHARMa::testDispersion(sim_size)
```

##### Estimated marginal means (back-transformed to mm)

```{r}
emm_size <- emmeans(m_size_tweedie, ~ season, type = "response")
emm_size
pairs(emm_size, adjust = "tukey")
```

##### Figure: model-based seasonal means

```{r}
df_size_emm <- as.data.frame(emm_size)

p_size_emm <- ggplot(df_size_emm, aes(x = season, y = response)) +
  ggdist::stat_slab(data = limpet_size, aes(x = season, y = sl_mm, fill = season),adjust = 0.7, width = 0.6, justification = -0.15, show.legend = FALSE)+
  geom_jitter(data = limpet_size, aes(x = season, y = sl_mm, color = season), width = 0.08, height = 0, alpha = 0.12, size = 0.8)+
  geom_point(size = 2) +
  geom_errorbar(data = df_size_emm, aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.12) +
   geom_signif(comparisons = list(c("Fall", "Spring")),
             map_signif_level = TRUE, # Automatically uses asterisks for p-values
             annotations = "**")+ # Manually set the annotation.
   geom_signif(comparisons = list(c("Winter", "Spring")),
              map_signif_level = TRUE, # Automatically uses asterisks for p-values
              annotations = "***")+
  scale_fill_manual(values = season_colors)+
  scale_color_manual(values =season_colors)+
  labs(x = NULL, y = "Estimated mean shell length (mm)",
       title = "Seasonal effect on owl limpet size (Tweedie GLMM)"
       ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  theme(legend.position = "none")
ggsave(filename = "figure4.png", width = 14, height = 6)

p_size_emm
```

### Size (`sl_mm`) by size (>30mm) by transect

Since shell length has a bimodal distribution, we decided that separating large and small would provide a a more complex story

```{r}
small_limpet_count <- limpet_size %>% 
  filter(size == "small") %>% 
  group_by(season) %>% 
  summarise(n_sl= n())

small_limpet_count

small_limpet_freq <- limpet_size %>% 
  filter(size == "small") %>%
  group_by(season) %>% 
  mutate(rounded_sl = round(sl_mm, 0)) %>% 
  count(rounded_sl)

p_small_size <- limpet_size %>%
  filter(size == "small") %>% 
  ggplot(aes(x = season, y = sl_mm, color = season)) +
  ggdist::stat_halfeye(adjust = 0.7, width = 0.6, justification = -0.15, aes(fill = season)) +
  geom_boxplot(width = 0.18, outlier.shape = NA) +
  geom_jitter(width = 0.08, height = 0, alpha = 0.20, size = 0.8) +
  labs(x = NULL, y = "Shell length (mm)",
       title = "Small Limpet Size Distribution per Season"
       ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  scale_color_manual(values = season_colors)+
  scale_fill_manual(values = season_colors)+
  theme(legend.position = "none")


p_small_size
```

#### GLMM to identify difference of shell length mean of small individuals per season)

```{r}
small_sl <- limpet_size %>% 
  filter(size == "small")

m_small_size_lmer <- lmer(
  sl_mm ~ season+(1|plot_name),  
  data = small_sl)

summary(m_small_size_lmer)
check_model(m_small_size_lmer)
```


##### Estimated marginal means for small individuals(back-transformed to mm)

```{r}
small_emm_size <- emmeans(m_small_size_lmer, ~ season, type = "response")
small_emm_size
pairs(small_emm_size, adjust = "tukey")
```
##### Figure: model-based seasonal means for small individuals

```{r}
df_small_size_emm <- as.data.frame(small_emm_size)

p_small_size_emm <- ggplot(df_small_size_emm, aes(x = season, y = emmean)) +
  ggdist::stat_slab(data = small_sl, aes(x = season, y = sl_mm, fill = season),adjust = 0.7, width = 0.6, justification = -0.15, show.legend = FALSE)+
  geom_jitter(data = small_sl, aes(x = season, y = sl_mm, color = season), width = 0.08, height = 0, alpha = 0.12, size = 0.8)+
  geom_point(size = 2) +
  geom_errorbar(data = df_small_size_emm, aes(ymin = lower.CL, ymax = upper.CL), width = 0.12) +
  scale_fill_manual(values = season_colors)+
  scale_color_manual(values =season_colors)+
  geom_signif(comparisons = list(c("Fall", "Spring")),
              map_signif_level = TRUE, # Automatically uses asterisks for p-values
              annotations = "**")+ # Manually set the annotation.
   geom_signif(comparisons = list(c("Winter", "Spring")),
              map_signif_level = TRUE, # Automatically uses asterisks for p-values
              annotations = "**")+
  labs(x = NULL, y = "Estimated mean shell length (mm)",
       title = "Seasonal effect on small owl limpet size (LMM)"
       ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  theme(legend.position = "none")
ggsave(filename = "p_small_size_emm.png", width = 14, height = 6)

p_small_size_emm

```
### Size (`sl_mm`) by transect for large limpets

Here I will analyze large individuals (>30 mm)

```{r}
large_limpet_count <- limpet_size %>% 
  filter(size == "large") %>% 
  group_by(season) %>% 
  summarise(n_sl= n())

large_limpet_count

large_limpet_freq <- limpet_size %>% 
  filter(size == "large") %>%
  group_by(season) %>% 
  mutate(rounded_sl = round(sl_mm, 0)) %>% 
  count(rounded_sl)


p_large_size <- limpet_size %>%
  filter(size == "large") %>% 
  ggplot(aes(x = season, y = sl_mm, color = season)) +
  ggdist::stat_halfeye(adjust = 0.7, width = 0.6, justification = -0.15, aes(fill = season)) +
  geom_boxplot(width = 0.18, outlier.shape = NA) +
  geom_jitter(width = 0.08, height = 0, alpha = 0.20, size = 0.8) +
  labs(x = NULL, y = "Shell length (mm)",
       title = "Large limpet size distribution by season"
       ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  scale_color_manual(values = season_colors)+
  scale_fill_manual(values = season_colors)+
  theme(legend.position = "none")


p_large_size
```

#### LMER for large individuals

```{r}
large_sl <- limpet_size %>% 
  filter(size == "large")

m_large_size_lmer <- lmer(
  sl_mm ~ season+(1|plot_name),  
  data = large_sl)

summary(m_large_size_lmer)
check_model(m_large_size_lmer)
```


##### Estimated marginal means for small individuals(back-transformed to mm)

```{r}
large_emm_size <- emmeans(m_large_size_lmer, ~ season, type = "response")
large_emm_size
pairs(large_emm_size, adjust = "tukey")
```
##### Figure: model-based seasonal means for large individuals

```{r}
df_large_size_emm <- as.data.frame(large_emm_size)

p_large_size_emm <- ggplot(df_large_size_emm, aes(x = season, y = emmean)) +
  ggdist::stat_slab(data = large_sl, aes(x = season, y = sl_mm, fill = season),adjust = 0.7, width = 0.6, justification = -0.15, show.legend = FALSE)+
  geom_jitter(data = large_sl, aes(x = season, y = sl_mm, color = season), width = 0.08, height = 0, alpha = 0.12, size = 0.8)+
  geom_point(size = 2) +
  geom_errorbar(data = df_large_size_emm, aes(ymin = lower.CL, ymax = upper.CL), width = 0.12) +
  scale_fill_manual(values = season_colors)+
  scale_color_manual(values =season_colors)+
  geom_signif(comparisons = list(c("Fall", "Spring")),
             map_signif_level = TRUE, # Automatically uses asterisks for p-values
             annotations = "**")+ # Manually set the annotation.
   geom_signif(comparisons = list(c("Winter", "Spring")),
              map_signif_level = TRUE, # Automatically uses asterisks for p-values
              annotations = "***")+
  labs(x = NULL, y = "Estimated mean shell length (mm)",
       title = "Transect effect on large owl limpet size (GLMM)"
       ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  theme(legend.position = "none")
ggsave(filename = "p_large_size_emm.png", width = 14, height = 6)

p_large_size_emm
```

```{r}
season_size_transect <- limpet_size %>%
  ggplot(aes(x = size, y = sl_mm)) +
  geom_boxplot(width = 0.3, outlier.shape = NA) +
  geom_jitter(aes(color = size),width = 0.1, height = 0, alpha = 0.10, size = 0.8) +
  labs(x = NULL, y = "Shell length (mm)",
       title = "Owl limpet size distribution by season"
       ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
   scale_color_manual(values = size_colors)+
  theme(legend.position = "none")+
  facet_wrap(~season)


season_size_transect
```

```{r}
p_size_season <- limpet_size %>%
  ggplot(aes(x = size, y = sl_mm)) +
  ggdist::stat_halfeye(adjust = 0.7, width = 0.6, justification = -0.15, alpha = 0.80, aes(fill = size)) +
  geom_boxplot(width = 0.18, outlier.shape = NA) +
  geom_jitter(aes(color = size),width = 0.08, height = 0, alpha = 0.08, size = 0.8) + #not sure if i should color the jitter 
  labs(x = NULL, y = "Shell length (mm)",
       title = "Owl limpet size distribution by season"
       ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  scale_color_manual(values = size_colors)+
  scale_fill_manual(values = size_colors)+
  theme(legend.position = "none")+
  facet_wrap(~season)


p_size_season
```

```{r}
#| fig-width: 14
#| fig-height: 6
## Bring the figures together for pub quality
## Add in stars with brackes to show significance between seasons based on emmeans
small_large_SL <- (p_small_size_emm+labs(title = NULL)|p_large_size_emm+labs(title = NULL))+plot_annotation(tag_levels = "a")
small_large_SL
ggsave(filename = "small_large_sl.png", width = 14, height = 6)
#title = "Seasonal effects on Owl Limpet Density and Shell Length"
```

```{r}
size_season_gamma <- glmmTMB(
  sl_mm ~ size*season + (1 | plot_name),
  family = Gamma(link = "log"),
  dispformula = ~ season,  # Model dispersion as a function of season because unequal variances
  data = limpet_size)

check_model(size_season_gamma)
summary(size_season_gamma)
```

##### Diagnostics

```{r}
size_seas_sim <- DHARMa::simulateResiduals(size_season_gamma, n = 1000)
plot(size_seas_sim)
DHARMa::testDispersion(size_seas_sim)
```

##### Estimated marginal means for size and season(back-transformed to mm)

```{r}
size_season_emm <- emmeans(size_season_gamma, ~ size*season, type = "response")
size_season_emm
pairs(size_season_emm, adjust = "tukey")
```

##### Figure: model-based size means by season

```{r}
df_size_season_emm <- as.data.frame(size_season_emm)

p_size_season_emm <- ggplot(df_size_season_emm, aes(x = size, y = response)) +
  ggdist::stat_slab(data = limpet_size, aes(x = size, y = sl_mm, fill = size),adjust = 0.7, width = 0.6, justification = -0.15, alpha = 0.8, show.legend = FALSE)+
  geom_jitter(data = limpet_size, aes(x = size, y = sl_mm, color = size), width = 0.08, height = 0, alpha = 0.08, size = 0.8)+
  geom_point(size = 2) +
  geom_errorbar(data = df_size_season_emm, aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.12) +
  scale_fill_manual(values = size_colors)+
  scale_color_manual(values =size_colors)+
  #geom_signif(comparisons = list(c("Fall", "Spring")),
             # map_signif_level = TRUE, # Automatically uses asterisks for p-values
              #annotations = "**")+ # Manually set the annotation.
   #geom_signif(comparisons = list(c("Winter", "Spring")),
              #map_signif_level = TRUE, # Automatically uses asterisks for p-values
              #annotations = "***")+
  labs(x = NULL, y = "Estimated mean shell length (mm)",
       title = "Size effect on owl limpet shell length by season (Gamma GLMM)"
       ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.08)))+
  theme(legend.position = "none")+
  facet_wrap(~season)
ggsave(filename = "size_seasons_SL.png", width = 14, height = 6)

p_size_season_emm

```


#### Different kind of visual to look at size structure shifts (CDFs)

Cumulative distributions are great for communicating shifts in size structure (e.g., truncation).

You would interpret this plot by saying: “For each season, what fraction of limpets are smaller than a given shell length?” 

Example interpretations of the plot

Steeper ECDF at small sizes in Spring
→ Maybe recent recruitment or better post-settlement survival

Right-shifted ECDF in Fall/Summer 
→ growth of survivors after spring recruitment?

These patterns often cannot be detected by mean-based models alone.

“We visualized seasonal shifts in owl limpet size structure using empirical cumulative distribution functions (ECDFs), which depict the proportion of individuals that are smaller than a given shell length without assuming an underlying distribution.”
```{r}
p_cdf <- limpet_size %>%
  ggplot(aes(x = sl_mm, color = season)) +
  stat_ecdf(linewidth = 1) +
  scale_color_manual(values = season_colors)+
  labs(x = "Shell length (mm)", y = "Empirical CDF",
       title = "Seasonal shifts in size structure (ECDF)", color = "Season")
ggsave(filename="cdf.png", width = 10, height = 6)

p_cdf
```

---

You would start with the mean values that you get from the emmeans output of the size model, state if there are any differences and then interpretit this model. The empirical cumulative distribution functions (ECDFs) showed  seasonal differences in size structure beyond mean responses. For a given shell length, the proportion of individuals by size differed among seasons, indicating seasonal shifts in the relative abundance of small versus large individuals. In particular, Spring exhibited a higher proportion of small individuals across much of the size range, whereas Fall and Winter showed a right-shifted distribution consistent with a greater representation of larger individuals.


```{r}
#| fig-width: 14
#| fig-height: 6
## Bring the figures together for pub quality
## Add in stars with brackes to show significance between seasons based on emmeans
OL_den_sl <- (p_density_emm+labs(title = NULL)|p_size+labs(title = NULL)|p_cdf+labs(title = NULL))+plot_annotation(tag_levels = "a")
OL_den_sl
ggsave(filename = "figure1.png", width = 14, height = 6)
#title = "Seasonal effects on Owl Limpet Density and Shell Length"
```

##### Physical Metrics
Because my work is heavily tied to physical changes that is correlated to seasonal change, I have pulled temperature and upwelling data from NOAA's ERDAP data. The upwelling data that I used is the CUTI(Coastal Upwelling Transport Index). 
Temperature data is sea surface temperature that is collected from buoys.

```{r}
# CUTI Data
# This pulls data
  cnm_upwelling_data<-griddap(datasetx = "erdCUTIdaily", 
                         url = "https://oceanview.pfeg.noaa.gov/erddap/", 
                         time = c("2023-06-15", "2024-06-15"),
                         latitude =  c(32.0, 33.0),
                         fields = "CUTI")$data %>% 
    mutate(date = as.Date(stringr::str_remove(time, "T00:00:00Z"))) %>% 
    select(latitude, date, CUTI) %>% 
    na.omit()%>%
    mutate(month = month(date), # extract the month
           year = year(date),
           site = "CNM Zone III",
           season  = case_when( # make a season
             month %in% c(12,1,2) ~ "Winter",
             month %in% c(3,4,5) ~ "Spring",
             month %in% c(6,7,8) ~ "Summer",
             month %in% c(9,10,11) ~ "Fall"
           ),
           season = factor(season),
           season = fct_relevel(season, c("Fall","Winter","Spring","Summer")))
  
upw_line <- ggplot(cnm_upwelling_data, aes(x = date, y = CUTI))+
    geom_point(aes(color = season, alpha = 2))+
    guides(alpha = "none")+
    geom_smooth(method = "loess", se = TRUE, color = "black")+
  scale_color_manual(values = season_colors)+
    labs(x=NULL, y= "Coastal Upwelling Transport Index", color = "Season")+
    theme_bw()

upw_line    
  
upw_box <- cnm_upwelling_data %>%
    filter(latitude== "32") %>% 
    ggplot(aes(x = as.factor(year), y = CUTI, fill = season))+
    geom_boxplot()+
  scale_fill_manual(values = season_colors)+
    labs(fill = "Season",
         x = "Year",
         y = "Upwelling")
upw_box

##### Temperature Data
# This Pulls Data
cnm_bouy_temp_data<-griddap(datasetx = "ncdcOisst21Agg_LonPM180", 
                   url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                   time = c("2023-06-15", "2024-06-15"), 
                   zlev = c(0,0), # this is the depth, it wants a range
                   latitude =  c(32.375,32.875),
                   longitude = c(-117.625,-117.625),
                   fields = "sst")$data %>% 
  mutate(time = as.Date(stringr::str_remove(time, "T00:00:00Z"))) %>% 
  select(longitude, latitude, time, sst) %>% 
  na.omit()%>%
  mutate(month = month(time), # extract the month
         year = year(time),
         site = "Cabrillo National Monument",
         season  = case_when( # make a season
           month %in% c(12,1,2) ~ "Winter",
             month %in% c(3,4,5) ~ "Spring",
             month %in% c(6,7,8) ~ "Summer",
             month %in% c(9,10,11) ~ "Fall"
         ),
         season = factor(season),
         season = fct_relevel(season, c("Fall","Winter","Spring","Summer")))

temp_line <- ggplot(cnm_bouy_temp_data, aes(x = time, y = sst))+
  geom_point(aes(color = season, alpha = 2))+
    guides(alpha = "none")+
    guides(color = "none")+
    geom_smooth(method = "loess", se = TRUE, color = "black")+
  scale_color_manual(values = season_colors)+
    labs(x=NULL, y= "Sea Surface Temperature (C°)", color = "Season")+
    theme_bw()
temp_line

temp_box <- cnm_bouy_temp_data %>%
    ggplot(aes(x = as.factor(year), y = sst, fill = season))+
    geom_boxplot()+
  scale_fill_manual(values = season_colors)+
    labs(fill = "Season",
         x = "Year",
         y = "Sea Surface Temperature (C°)")
temp_box

```
```{r}
#| fig-width: 14
#| fig-height: 6
## Bring the figures together for pub quality
upline_templine_cnm <- (temp_line+labs(title = NULL) | upw_line+labs(title = NULL)+plot_annotation(tag_levels = "a"))
upline_templine_cnm
ggsave(filename = "upline_templine_cnm.png", width = 14, height = 6)
#title = "Seasonal effects on Owl Limpet Density and Shell Length"
```


#### Site Map
```{r}
zone3 <- data.frame(lon = -117.2444128, lat = 32.666561)

map1 <- get_map(zone3, zoom = 19, , maptype = "satellite")

site_map <- ggmap(map1)+
  geom_point(data = transect_coordinates, 
             aes(x = long, y = lat, color = transect))
site_map
ggsave(filename = "site_map.png", width = 14, height = 6)

```



## Question 2: Does limpet presence affect community composition, and does it interact with season?

We analyze both:
- **Mobile invertebrate counts** (multivariate & univariate)
- **Sessile cover composition** (multivariate & univariate) --  I am having a hard time with the sessile data. Seems like everything is mostly bare rock and I am not sure what the microalgae are here

Key fixed effects:
- `col_pres` (No vs Yes)
- `season`
- `col_pres * season` interaction

Random/blocking structure:
- `transect` is used for random effects and permutation strata (to respect spatial non-independence).

---


## 2A. Mobile invertebrates (counts)

### Identify species columns

```{r}
# group by species groups first

mobile_grouped<-mobile %>%
  mutate(Limpets = rowSums(across(c(l_limatula, l_scabra, l_gigantea))),
         Chitons = rowSums(across(c(nuttallina_sp, mopalia_sp, l_hartwegii,l_dentiens))),
         Turbans = rowSums(across(c(t_eiseni:t_gallina))),
         Periwinkles = rowSums(across(c(l_plena, l_keenae))),
         Whelks = m_lugubris) %>%
  select(-c(m_lugubris:l_gigantea)) # what is -c

# metadata
mobile_meta <- c("season","date","transect","plot_name","l_r","meter_no","col_pres","t_f","tag_id")
mobile_spp <- mobile_grouped %>% select(-any_of(mobile_meta)) #what is -any_of

# Sanity checks: are these non-negative counts?
summary(mobile_spp)
```

### Transformations for multivariate analyses

For community composition, we typically:
- use **gower** on square-root transformed counts since lots of zeros. Bray-Curtis did not work.

For Gower, multiplying all counts by the same constant (e.g., converting to per m² when all plots have identical area) does **not** change dissimilarities; we still report univariate abundance in **m⁻²** for interpretability.
Information on coding dissimilarity indices https://rdrr.io/cran/vegan/man/vegdist.html


```{r}
mobile_mat <- mobile_spp %>%
  mutate(across(everything(), ~replace_na(., 0))) %>% # make the missing values zeros
  as.matrix()

# gower dissimilarity on sqrt counts because 0s present
mobile_gower <- vegdist(sqrt(mobile_mat), method = "gower")
```

### PERMANOVA (adonis2) with permutations constrained by transect

This tests whether multivariate centroids differ by predictors.

```{r}
set.seed(1)


adon_mobile <- adonis2(
  mobile_gower ~ col_pres * season,
  data = mobile,
  permutations = 999,
  strata = mobile$transect
)



print(adon_mobile)
pairwise.adonis2(x=mobile_gower ~ col_pres * season,
                 data = mobile,
                 method = "gower",
                 p.adjust.m = "bonferroni")
## Need to add in the pairwise differences-- follow up with this
#What about pairwise.adonis2()
```

### Check dispersion (PERMDISP)

PERMANOVA can be sensitive to differences in within-group dispersion.

```{r}
disp_mobile <- betadisper(mobile_gower, interaction(mobile$col_pres, mobile$season))
anova(disp_mobile)
permutest(disp_mobile, permutations = 999)
tukey_disp <- TukeyHSD(disp_mobile) # get pairwise differences
tukey_disp
# Yes clearly very different dispersion (the size of the circle)--but driven by season based on pairwise
# Yes.Fall-No.Fall p=0.5271168
#Yes.Winter-No.Winter p =0.4522087
#Yes.Spring-No.Spring p = 0.9995178


```

### Ordination: NMDS (visualize patterns)
Website with info on metaMDS https://vegandevs.github.io/vegan/reference/metaMDS.html
```{r}
set.seed(2)

nmds_mobile <- metaMDS(sqrt(mobile_mat), distance = "gower", k = 2, trymax = 100)

nmds_mobile$stress #get the stress
```


```{r}
# --- NMDS biplot for MOBILE invertebrates ( species vectors) ----


set.seed(42) 

# Fit species vectors to the NMDS using envfit. envfit finds vectors or vector averages
#    - "display = 'species'" returns directions for each taxa
#    - we fit to sqrt(mobile_mat) if NMDS was based on sqrt counts
fit_spp <- envfit(nmds_mobile, sqrt(mobile_mat), permutations = 999)

# Extract site scores
sites_df <- scores(nmds_mobile, display = "sites") %>%
  as.data.frame() %>%
  rownames_to_column("site_id") %>%
  bind_cols(mobile_grouped %>% mutate(site_id = row_number()) %>% select(site_id, season, col_pres, transect))

# Extract species vectors + p-values; keep only significant / strong ones
vec_df <- scores(fit_spp, display = "vectors") %>%
  as.data.frame() %>%
  rownames_to_column("taxon") %>%
  # p-values are in fit_spp$vectors$pvals; same order as rows in scores()
  mutate(p = fit_spp$vectors$pvals) %>%
  # vector strength (longer = stronger relationship with ordination)
  mutate(r = sqrt(NMDS1^2 + NMDS2^2)) %>%
  arrange(p) ## all the species have significant vectors

# Choose which taxa to label: either significant + top N by vector length
alpha <- 0.05
top_n <- 12

vec_keep <- vec_df %>%
  filter(p <= alpha) %>%
  slice_max(order_by = r, n = top_n)

# Scale vectors to match plot extent (so arrows look nice)
#    This rescales arrows relative to NMDS axes range.
arrow_mult <- 0.85 * min(
  diff(range(sites_df$NMDS1)) / diff(range(vec_keep$NMDS1)),
  diff(range(sites_df$NMDS2)) / diff(range(vec_keep$NMDS2))
)

vec_keep <- vec_keep %>%
  mutate(NMDS1 = NMDS1 * arrow_mult,
         NMDS2 = NMDS2 * arrow_mult)

# Plot: sites + ellipses + species vectors (biplot)
p_biplot_mobile <- ggplot(sites_df, aes(NMDS1, NMDS2)) +
  # Sites
  geom_point(aes(color = season, shape = col_pres), alpha = 0.85, size = 2) +
  # Group ellipses
  stat_ellipse(aes(group = interaction(season, col_pres), color = season),
               level = 0.68, linewidth = 0.7, show.legend = FALSE) +
  # Species arrows
  geom_segment(
    data = vec_keep,
    aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
    inherit.aes = FALSE,
    linewidth = 0.7,
    arrow = arrow(length = unit(0.18, "cm"))
  ) +
  # Species labels
  ggrepel::geom_text_repel(
    data = vec_keep,
    aes(x = NMDS1, y = NMDS2, label = taxon),
    inherit.aes = FALSE,
    size = 3,
    max.overlaps = Inf,
    min.segment.length = 0
  ) +
  scale_color_manual(values = season_colors) +
  labs(
    title = "Mobile invertebrates NMDS biplot",
    subtitle = paste0("Stress = ", round(nmds_mobile$stress, 3)),
    color = "Season",
    shape = "Owl limpet present?"
  ) +
  theme_classic()+
  theme(legend.position = "none")

p_biplot_mobile

```


### Univariate summaries (total abundance)

These are interpretable “whole-community” metrics.

```{r}
mobile_alpha <- mobile_grouped %>%
  rowwise() %>%
  mutate(
    #  mobile plot is 0.25 m² -- This is right
    area_m2 = ifelse(season == "Fall", 0.2, 0.25), #fall plots are circular, this is area of a circle
    total_abund = sum(c_across(all_of(colnames(mobile_mat))), na.rm = TRUE),
    total_abund_m2 = total_abund / area_m2,
  ) %>%
  ungroup()

mobile_alpha %>% select(season, col_pres, transect, total_abund) %>% glimpse()
```

#### Mixed models for Individual metrics

- `total_abund`: NB GLMM

```{r}
m_mobile_total <- glmmTMB(
  # Model counts with area offset to infer abundance density (per m²)
  total_abund ~ col_pres * season + offset(log(area_m2)) + (1 | transect),
  family = nbinom2(),
  data = mobile_alpha)


summary(m_mobile_total)
emm_total <- emmeans(m_mobile_total, ~ col_pres|season, type = "response")
pairs(emm_total, adjust = "tukey")

performance::check_model(m_mobile_total)

# Species pairs
# Limpets
m_mobile_Limpet <- glmmTMB(
  # Model counts with area offset to infer abundance density (per m²)
  Limpets ~ col_pres * season + offset(log(area_m2)) + (1 | transect),
  family = nbinom2(),
  data = mobile_alpha)


summary(m_mobile_Limpet)
emm_limpet <- emmeans(m_mobile_Limpet, ~ col_pres|season, type = "response")
pairs(emm_limpet, adjust = "tukey")
limpet_pairs <- pairs(emm_limpet, adjust = "tukey")
limpet_emm_pairs_df <- as.data.frame(limpet_pairs, adjust = "tukey")
performance::check_model(m_mobile_Limpet)

# Periwinkles
m_mobile_Periwinkles <- glmmTMB(
  # Model counts with area offset to infer abundance density (per m²)
  Periwinkles ~ col_pres * season + offset(log(area_m2)) + (1 | transect),
  family = nbinom2(),
  data = mobile_alpha)


summary(m_mobile_Periwinkles)
emm_Periwinkles <- emmeans(m_mobile_Periwinkles, ~ col_pres|season, type = "response")
pairs(emm_Periwinkles, adjust = "tukey")
periwinkle_pairs <- pairs(emm_Periwinkles, adjust = "tukey")
periwinkle_emm_pairs_df <- as.data.frame(periwinkle_pairs, adjust = "tukey")
performance::check_model(m_mobile_Periwinkles)

# Chitons
m_mobile_Chitons <- glmmTMB(
  # Model counts with area offset to infer abundance density (per m²)
  Chitons ~ col_pres * season + offset(log(area_m2)) + (1 | transect),
  family = nbinom2(),
  data = mobile_alpha)


summary(m_mobile_Chitons)
emm_Chitons <- emmeans(m_mobile_Chitons, ~ col_pres|season, type = "response")
pairs(emm_Chitons, adjust = "tukey")
chiton_pairs<- pairs(emm_Chitons, adjust = "tukey")
chiton_emm_pairs_df <- as.data.frame(chiton_pairs, adjust = "tukey")
performance::check_model(m_mobile_Chitons) #have high collineararity coefficient

# Turbans
m_mobile_Turbans <- glmmTMB(
  # Model counts with area offset to infer abundance density (per m²)
  Turbans ~ col_pres * season + offset(log(area_m2)) + (1 | transect),
  family = nbinom2(),
  data = mobile_alpha)


summary(m_mobile_Turbans)
emm_Turbans <- emmeans(m_mobile_Turbans, ~ col_pres|season, type = "response")
pairs(emm_Turbans, adjust = "tukey")
turban_pairs <- pairs(emm_Turbans, adjust = "tukey")
turban_emm_pairs_df <- as.data.frame(turban_pairs, adjust = "tukey")
performance::check_model(m_mobile_Turbans) #have high collineararity coefficient

# Whelks
m_mobile_Whelks <- glmmTMB(
  # Model counts with area offset to infer abundance density (per m²)
  Whelks ~ col_pres * season + offset(log(area_m2)) + (1 | transect),
  family = nbinom2(),
  data = mobile_alpha)


summary(m_mobile_Whelks)
emm_Whelks <- emmeans(m_mobile_Whelks, ~ col_pres|season, type = "response")
pairs(emm_Whelks, adjust = "tukey")
whelk_pairs <- pairs(emm_Whelks, adjust = "tukey")
whelk_emm_pairs_df <- as.data.frame(whelk_pairs, adjust = "tukey")
performance::check_model(m_mobile_Whelks) #have high collineararity coefficient

combined_contrast <- rbind((limpet_emm_pairs_df),(periwinkle_emm_pairs_df), (chiton_emm_pairs_df), (turban_emm_pairs_df), (turban_emm_pairs_df))
```

#### Figure: total abundance by COL presence and season

```{r}
#| fig-width: 14
#| fig-height: 6 

# everything
p_alpha <- mobile_alpha %>%
  pivot_longer(cols = c(Limpets:Whelks, total_abund_m2), 
               names_to = "metric", 
               values_to = "value") %>%
  mutate(value_log = value+1)%>%
  ggplot(aes(x = season, 
             y = value_log, 
             shape = col_pres, 
             color = season)) +
  coord_trans(y = "log")+
  scale_color_manual(values = season_colors)+
  scale_fill_manual(values = season_colors)+
 # ggdist::stat_halfeye(position = position_dodge(width = 0.7), alpha = 0.7) +
  geom_violin(position = position_dodge(width = 0.7), alpha = 0.25, linewidth = 0.5,
              aes(fill = season), color = NULL, show.legend = FALSE)+ #outlier.shape = NA, alpha = 0.5) +
  stat_summary(position = position_dodge(width = 0.7))+
  facet_wrap(~ metric, scales = "free_x")+
          scale_y_continuous(breaks = c(1,10,100,200))+
            labs(y = "Density (m⁻²)", x = NULL, fill = "Owl limpet present?", color = "Season", shape = "Present")+
      # title = "Mobile invertebrate alpha metrics vs limpet presence and season" 
  theme(legend.position = "top")

## total abundance only
p_total <- mobile_alpha %>%
  ggplot(aes(x = season, y = total_abund_m2+1, shape = col_pres, color = season)) +
  coord_trans(y = "log")+
  scale_color_manual(values = season_colors)+
  scale_fill_manual(values = season_colors)+
 # ggdist::stat_halfeye(position = position_dodge(width = 0.7), alpha = 0.7) +
  geom_violin(position = position_dodge(width = 0.7), alpha = 0.25, linewidth = 0.5,
              aes(fill = season), color = NULL, show.legend = FALSE)+ #outlier.shape = NA, alpha = 0.5) +
  stat_summary(position = position_dodge(width = 0.7), size = 1.5)+
   labs(y = "Total density (m⁻²)", x = NULL, shape = "Owl limpet present?",
      # title = "Mobile invertebrate alpha metrics vs limpet presence and season"
       ) +
  theme(legend.position = "top")


## add in brackets for yes/no pairs for ones that are significant based on emmeans and clean up the title of total_abunc... and double check that the area is correct since I think you did it two different ways
p_alpha+p_biplot_mobile+plot_annotation(tag_levels = "a")
ggsave(filename = "figure2.png", width = 14, height = 6)

#graph with brackets was saved to E:\Thesis Data\R Figures
```

```{r}
herb_guild <- mobile_spp %>%
  select(Limpets, Chitons,Turbans, Periwinkles) %>% 
  mutate(across(everything(), ~replace_na(., 0))) %>% # make the missing values zeros
  as.matrix()

select_mobiles <- mobile_alpha %>% 
  filter(col_pres == "No") %>% 
  select(season, transect, plot_name, col_pres, Limpets, Chitons, Turbans, Periwinkles) 

name_change_limpet_den <- limpet_density %>% 
  rename(transect = plot_name) %>% 
  select(season, transect, size, n_limpets, area_m2, density_m2, mean_sl, sd_sl)

limp_comm_join <- full_join(name_change_limpet_den, select_mobiles, by =c("season", "transect"))

limp_dens_per_tran <- limp_comm_join %>% 
  group_by(transect) %>% 
  rowwise() %>%
  mutate(area_all_cont = ifelse(season == "Fall", 1.4, 1.75), #fall plots are circular, this is area of a circle. I multiplied the area by 7 to get the total area of all of the controls per transect
          total_abund = sum(c_across(all_of(colnames(herb_guild))), na.rm = TRUE),
          total_abund_total_area = total_abund / area_all_cont,
  ) %>%
  ungroup()
  
```


#### Scatter Plot to visualize regression of all limpet density across transects v abundance of herbivores in no central limpet plots (controls)
```{r}
limp_v_herbvs_plot <- limp_dens_per_tran %>% 
  ggplot(aes(x = density_m2, y = total_abund_total_area, color= transect))+
  geom_point()+
  geom_smooth(method = "lm", se = TRUE, color = "blue")+
  facet_wrap(~season)+
  labs(x = "Owl Limpet Density", y = "Density of Competitors (total area)",
       title = "Linear Relationship Between Owl Limpet Density and Abundance of Competitors"
       )+
  theme_bw()

limp_v_herbvs_plot
```
#LMM for regression of all limpet density across transects v abundance of herbivores in no central limpet plots
```{r}
qqnorm(limp_dens_per_tran$total_abund_total_area)
total_den_limp_vherbs <- lmer(
  # Model counts with area offset to infer abundance density (per m²)
  total_abund_total_area ~ density_m2 * season + (1 | transect),
  data = limp_dens_per_tran)

slopes <- emtrends(total_den_limp_vherbs, pairwise ~ season, var = "density_m2")
summary(slopes)
check_model(total_den_limp_vherbs)
```

#### Scatter Plot to visualize regression for density of competitors in non limpet plots vs density of large limpets in transects
```{r}
large_limp_dens_per_tran <- limp_dens_per_tran %>% 
  filter(size == "large")%>% 
  group_by(transect, season)

large_limp_v_herbvs_plot <- large_limp_dens_per_tran %>%
  ggplot(aes(x = density_m2, y = total_abund_total_area, color = transect))+
  geom_point()+
  geom_smooth(method = "lm", se = TRUE, color = "blue")+
  facet_wrap(~season)+
  labs(x = "Owl Limpet Density", y = "Density of Competitors (total area)",
       title = "Linear Relationship Between Large Owl Limpet Density and Abundance of Competitors"
       )+
  theme_bw()

large_limp_v_herbvs_plot
```
#####LMM for regression for density of competitors in non limpet plots vs density of large limpets in transects
```{r}
large_den_limp_vherbs <- lmer(total_abund_total_area ~ density_m2 * season + (1 | transect),
  data = large_limp_dens_per_tran)

large_slopes <- emtrends(large_den_limp_vherbs, pairwise ~ season, var = "density_m2")
summary(large_slopes)
check_model(total_den_limp_vherbs)
```


#### Scatter Plot to visualize regression for density of competitors in non limpet plots vs density of small limpets in transects
```{r}
small_limp_v_herbvs <-limp_dens_per_tran %>%
  filter(size == "small") %>% 
  group_by(transect, season) 

small_limp_v_herbvs_plot <- small_limp_v_herbvs %>% 
  ggplot(aes(x = density_m2, y = total_abund_total_area, color = transect))+
  geom_point()+
  geom_smooth(method = "lm", se = TRUE, color = "blue")+
  facet_wrap(~season)+
  labs(x = "Owl Limpet Density", y = "Density of Competitors (total area)",
       title = "Linear Relationship Between Small Owl Limpet Density and Abundance of Competitors"
       )+
  theme_bw()

small_limp_v_herbvs_plot
```

```{r}
small_den_limp_vherbs <- lmer(total_abund_total_area ~ density_m2 * season + (1 | transect),
  data = small_limp_v_herbvs)

small_slopes <- emtrends(small_den_limp_vherbs, pairwise ~ season, var = "density_m2")
summary(small_slopes)
check_model(small_den_limp_vherbs)
```

---

## Sessile community (% cover / proportional cover)

### Identify sessile cover columns --- i dont trust this analysis rn

```{r}
sess_meta <- c("season","date","transect","plot_name","l_r","meter_no","col_pres","t_f","tag_id")
sess_spp <- sessile %>% select(-any_of(sess_meta))

summary(sess_spp)
```

### Multivariate test: PERMANOVA on gower


```{r}
sess_mat <- sess_spp %>%
  mutate(across(everything(), ~replace_na(., 0))) %>%
  as.matrix()

sess_bc <- vegdist(sqrt(sess_mat), method = "gower")

set.seed(3)
adon_sess <- adonis2(
  sess_bc ~ col_pres * season,
  data = sessile,
  permutations = 999,
  strata = sessile$transect
)

adon_sess
```

### Dispersion check

```{r}
disp_sess <- betadisper(sess_bc, interaction(sessile$col_pres, sessile$season))
anova(disp_sess)
permutest(disp_sess, permutations = 999)
```

### Ordination: NMDS

```{r}
set.seed(4)
nmds_sess <- metaMDS(sess_mat, distance = "gower", k = 2, trymax = 100)
nmds_sess$stress

scores_sess <- scores(nmds_sess, display = "sites") %>%
  as.data.frame() %>%
  bind_cols(sessile %>% select(season, col_pres, transect))

p_nmds_sess <- ggplot(scores_sess, aes(x = NMDS1, y = NMDS2, shape = col_pres, color = season)) +
  geom_point(alpha = 0.85, size = 2) +
  stat_ellipse(aes(group = interaction(season, col_pres)), level = 0.68, linewidth = 0.8) +
  labs(title = "Sessile community: NMDS (gower on proportional cover)",
       subtitle = paste0("Stress = ", round(nmds_sess$stress, 3)),
       shape = "Owl limpet present?", color = "Season") +
  theme(legend.position = "right")

p_nmds_sess
```

### Univariate analyses for key functional groups

Depending on your ecological framing,  response variables may include:
- **Bare rock** (space availability)
- **Erect fleshy algae** (here likely `ulva_sp`, `endocaldia_sp`, etc.)
- **Crustose coralline algae** (`coralline_crust`)
- **Barnacles** (`chtalmus_balanus`)
- **Tube worms** (`phragmatopoma_californica`) etc.

Where is the microalgae?

Because cover is proportional (0–1), a practical approach is:
- transform with `logit01()` and use LMM/GLMM with Gaussian errors, OR
- Beta regression (more complex; `glmmTMB` can do it with `family = beta_family()`).

Here we demonstrate logit-transformed LMMs with transect random intercepts.

```{r}
# Choose a few focal cover variables to test (edit for your hypotheses)
focal_cover <- c("bare_rock", "ulva_sp", "coralline_crust", "chtalmus_balanus", "ralfsia_sp")

sess_long <- sessile %>%
  pivot_longer(cols = all_of(focal_cover), names_to = "taxon", values_to = "cover") %>%
  mutate(
    cover = as.numeric(cover),
    cover_logit = logit01(cover)
  )

# Fit one model per taxon (pre-specified list!)
models_cover <- sess_long %>%
  group_by(taxon) %>%
  group_map(~{
    lmer(cover_logit ~ col_pres * season + (1 | transect), data = .x)
  })

names(models_cover) <- unique(sess_long$taxon)

# Summaries
lapply(models_cover, summary)
```

#### Figure: focal cover responses

```{r}
p_cover <- sess_long %>%
  ggplot(aes(x = season, y = cover, fill = col_pres)) +
#  geom_point(position = position_dodge(width = 0.7))+
 # ggdist::stat_halfeye(position = position_dodge(width = 0.7), alpha = 0.7) +
  geom_violin(position = position_dodge(width = 0.7), 
              # outlier.shape = NA, 
               alpha = 0.5) +
  facet_wrap(~ taxon, scales = "free_y", ncol = 2) +
  labs(x = NULL, y = "Proportional cover", fill = "Owl limpet present?",
       title = "Sessile functional/taxon cover vs limpet presence and season") +
  theme(legend.position = "top")

p_cover
```

---


## Ecological interpretation (how to read results -- fill in below with your references)

### 1) Seasonal density and size patterns (what’s plausible)

Depending on your results, common mechanisms include:

- **Recruitment pulses** (seasonal larval supply) can increase density of **small** limpets after settlement.
-- What else do you think?

For owl limpets specifically:
- *Lottia gigantea* is well known for **territorial grazing** that can maintain “gardens” of microalgae and influence space availability and algal cover (citations).  
- What else?

### 2) Limpet presence effects on community composition (expected directions)

- fill in


**Season interactions** are common because:
- why?

---

### potenitial storyline

**“Seasonal dynamics modulate the engineering role of a territorial grazer.”**

### Journals to think about

- *Marine Ecology Progress Series* (MEPS)
- *Journal of Experimental Marine Biology and Ecology*
- *PeerJ*

---


