---
title: "Community Surveys Visualizations"
author: "Cindy Fajardo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../Output/")
```

#Welcome!
In this repository I will have code to help visualize the raw community survey data collected at Cabrillo National Monument in San Diego, CA. I collected both sessile (percent cover) and mobile species (counts) data in the community found around the limpet population I am following for a year beginning September 2023. 

#Libraries
```{r, Libraries}
library("rerddap")
library(tidyverse)
library("tidync")
library("doParallel")
library(lubridate)
library(patchwork)
library(viridis)
library(here)
library(kableExtra)
library(hrbrthemes)
library(emmeans)
library(agricolae)
library(vegan)
library(performance)
library(broom)
library(car)
library(lme4)
library(lmerTest)
library(pscl)
library(ggridges)
library(devtools)
#library(pairwiseAdonis)
library(ggpubr)
library(glmmTMB)
library(RColorBrewer)
library(MASS)
library(DHARMa)
```

#Read in CSV
```{r, Read in CSV}
#feb_cnm_sessile <- read_csv(here("Data", "2024_02_CNM_Sessile_Data.csv"))
#feb_cnm_mobiles <- read_csv(here("Data","2024_02_CNM_Mobile_Data.csv"))
#feb_cnm_mob_count_mat <- read_csv(here("Data", "2024_02_CNM_Mobile_Count_Matrix.csv"))
#feb_cnm_sessile_count_mat <- read_csv(here("Data", "2024_02_CNM_Sessile_Count_Matrix.csv"))
#sessile_master <- read_csv(here("Data", "2023_09_CNM_Sessile_Data.csv"))
sess_percentcover <- read_csv(here("Data", "Master_CNM_Sessile_Percent_Data.csv"))
mobiles_master <- read.csv(here("Data", "Master_CNM_Mobile_Data.csv"))
mobiles_matrix_master <- read.csv(here("Data", "Master_CNM_Mobile_Matrix.csv"))
sessile_matrix_master <- read.csv(here("Data", "Master_CNM_Sessile_Matrix.csv"))
```

#Data Manipulation of Mobile Species
In this chunk, I manipulate the data to help visualize it
```{r, #Data Manipulation}
#feb_cnm_mobiles_sum_spec <- feb_cnm_mobiles %>% 
#  mutate(spec_sum = rowSums(across(c(M_lugubris, L_dentiens, L_plena, L_keenae, L_limatula, L_scabra, Nuttallina_sp, Mopalia_sp, P_crassipes, T_eiseni, T_funebralis, T_gallina, L_gigantea)), na.rm=TRUE)) %>% 
  # pivot_longer(cols = M_lugubris:L_gigantea, #pivoted to make longer
               #names_to = "Species", #all species name in this column
               #values_to = "Count")
```


#Boxplot of Mobile Organism Abundance Per Plot in Presence/Absence of Central Limpet
```{r,Boxplot of Mobile Organism Abundance Per Plot in Presence/Absence of Central Limpet}
 #feb_cnm_mobiles_sum_spec %>% 
  #ggplot(aes(
   # x = COL_pres,
    #y = spec_sum,
    #fill = COL_pres))+
  #geom_boxplot(show.legend = FALSE)+ 
 # scale_fill_brewer(palette = "Dark2")+
  #labs(x = "Presence of Central Owl Limpet") +
  #theme_bw()
  
```

#Data Manipulation of Sessile Species
Here, I manipulate the data to visualize the data. I separate it by animals and algae. 
```{r, #Data Manipulation of Sessile Species}
#feb_cnm_sessile_sum_plot <- feb_cnm_sessile %>%
 # mutate(spec_sum = rowSums(across(c(A_elegantissima, Phragmatopoma_californica, Balanus_Chtalmus, Bare_rock, Ralfsia_sp, Coralline_crust, Endocladia_sp, P_polymerus,Ulva_sp)), na.rm=TRUE))%>% 
  #mutate(Animal = (A_elegantissima+ Phragmatopoma_californica+ Balanus_Chtalmus+ P_polymerus)/spec_sum*100,
   #      Algae = (Coralline_crust+Endocladia_sp+Ulva_sp)/spec_sum*100)
```

#Boxplot of Sessile Organism Percent Cover Per Plot in Presence/Absence of Central Limpet
```{r, Boxplot of Sessile Organism Percent Cover Per Plot in Presence/Absence of Central Limpet}
#feb_cnm_sessile_sum_plot %>% 
 # pivot_longer(cols = c(Animal, Algae), names_to = "Percent_Type")%>%
  #ggplot(aes(
   # x = COL_pres,
    #y = value,
    #fill = Percent_Type))+
  #geom_boxplot(show.legend = TRUE)+ 
  #scale_fill_brewer(palette = "Pastel2")+
  #scale_color_brewer(palette = "Pastel1")+
  #labs(x = "Presence of Central Owl Limpet",
   #    fill = "Percent Type") +
  #theme_bw()

```

#Data Manipulation for Species Richness
```{r, Data Manipulation for Species Richness}
#feb_cnm_mob_count_mat$COL_pres <- as.factor(feb_cnm_mob_count_mat$COL_pres) #made this coulmn a factor
#feb_cnm_mob_count_mat$Transect<- as.factor(feb_cnm_mob_count_mat$Transect)

#USE THIS ONE
#feb_cnm_mob_count_mat_sum <- feb_cnm_mob_count_mat %>% 
 # mutate(spec_sum = rowSums(across(c(M_lugubris, L_dentiens, L_plena, L_keenae, L_limatula, L_scabra, Nuttallina_sp, Mopalia_sp, P_crassipes, T_eiseni, T_funebralis, T_gallina, L_gigantea)), na.rm=TRUE))

```

#Species Richness of Mobiles Boxplot
```{r, Species Richness of Mobiles Boxplot}
#feb_mobile_spec_rich_box <- feb_cnm_mob_count_mat_sum %>% 
 # ggplot(aes(
  #  x = COL_pres,
   # y = spec_sum,
    #fill = COL_pres))+
  #geom_boxplot(show.legend = FALSE)+ 
  #scale_fill_brewer(palette = "Dark2")+
  #labs(x = "Presence")+
  #theme_bw()
#feb_mobile_spec_rich_box
```

#Sessile Species Richnnes Data Manipulation 
```{r, Sessile Species Richnnes Data Manipulation}
#feb_cnm_sessile_count_mat$COL_pres <- as.factor(feb_cnm_sessile_count_mat$COL_pres) #made this coulmn a factor
#feb_cnm_sessile_count_mat$Transect<- as.factor(feb_cnm_sessile_count_mat$Transect)

#Using this one
#feb_cnm_sessile_count_mat_sum <- feb_cnm_sessile_count_mat %>% 
 #   select(Transect, Plot_Name, L_R, Meter_no, COL_pres, Tag_ID, Anthopleura_elegantissima,Phragmatopoma_californica, Balanus_glandula, Chthalmus, Brown_encrusting, Coralline_crust, Endocladia_sp, Goose_neck, Ulva ) %>% 
  #mutate(spec_sum = rowSums(across(c(Anthopleura_elegantissima,Phragmatopoma_californica, Balanus_glandula, Chthalmus, Brown_encrusting, Coralline_crust, Endocladia_sp, Goose_neck, Ulva)), na.rm=TRUE))

####Try this###
#feb_cnm_sessile_matrix <- feb_cnm_sessile_count_mat %>%
 # mutate(spec_sum = rowSums(across(c(Anthopleura_elegantissima,Phragmatopoma_californica, Balanus_glandula, Chthalmus, Bare_rock, Brown_encrusting, Coralline_crust, Endocladia_sp, Goose_neck, Ulva)), na.rm=TRUE))%>% 
  #mutate(Animal = (Anthopleura_elegantissima+ Phragmatopoma_californica+ Balanus_glandula+ Goose_neck+ Chthalmus),
   #      Algae = (Brown_encrusting+Coralline_crust+Endocladia_sp+Ulva))

```

#Boxplot of Sessile Species Richness
```{r, Boxplot of Sessile Species Richness}
#feb_cnm_sessile_matrix %>% 
 # pivot_longer(cols = c(Animal, Algae), names_to = "sessile_type")%>%
  #ggplot(aes(
   # x = COL_pres,
    #y = value,
    #fill = sessile_type))+
  #geom_boxplot(show.legend = TRUE)+ 
  #scale_fill_brewer(palette = "Pastel1")+
  #scale_color_brewer(palette = "Set1")+
  #labs(x = "Presence of Central Owl Limpet",
   #    fill = "Percent Type") +
  #theme_bw()

```

#Patchwork of Mobile Abundance and Sessile Percent Cover
```{r, Patchwork of Mobile Abundance and Sessile Percent Cover}
 #patch_mob <- feb_cnm_mobiles_sum_spec %>% 
  #ggplot(aes(
   # x = COL_pres,
    #y = spec_sum,
    #fill = COL_pres))+
  #geom_boxplot(show.legend = FALSE)+ 
  #scale_fill_brewer(palette = "Dark2")+
  #labs(title = "Abundance of Mobile Species",
   #    x = "Presence of Central Owl Limpet",
    #   y = "Abundance of Other Mobile Organisms") +
  #theme_bw()

#patch_sess <- feb_cnm_sessile_sum_plot %>% 
 # pivot_longer(cols = c(Animal, Algae), names_to = "percent_type")%>%
  #ggplot(aes(
   # x = COL_pres,
    #y = value,
    #fill = COL_pres,
    #color = percent_type))+
  #geom_boxplot(show.legend = TRUE)+ 
  #scale_fill_brewer(palette = "Pastel1")+
  #scale_color_brewer(palette = "Set1")+
  #labs(title = "Percent Cover of Sessile Species",
   #    x = "Presence of Central Owl Limpet",
    #   y = "Percent Cover") +
  #theme_bw()

#patch_mob+patch_sess+
 # plot_annotation(title = 'Winter Total Aundance of Mobiles Species and Percent Cover of Sessile Species in the Presence/Absence of Owl Limpets', theme = theme(plot.title = element_text(size = 12)))
```

#Patchwork of species richness for mobiles and sessile
```{r, Patchwork of species richness for mobiles and sessile}
#mobile_spec_rich_box_nolab <- feb_cnm_mob_count_mat_sum %>% 
 # ggplot(aes(
  #  x = COL_pres,
   # y = spec_sum,
    #fill = COL_pres))+
  #geom_boxplot(show.legend = FALSE)+ 
  #scale_fill_brewer(palette = "Dark2")+
  #labs(title = "Mobiles",
   #    x = "Presence/Absence",
    #   y = "Species Richness")+
  #theme_bw()

#sessile_spec_rich_nolab <- feb_cnm_sessile_count_mat_sum %>% 
 # ggplot(aes(
  #  x = COL_pres,
   # y = spec_sum,
    #fill = COL_pres))+
  #geom_boxplot(show.legend = FALSE)+
  #scale_fill_brewer(palette = "Dark2")+
  #labs(title = "Sessile",
   #    x = "Presence/Absence",
    #   y = "Species Richness")+
  #theme_bw()


#mobile_spec_rich_box_nolab+sessile_spec_rich_nolab+
 # plot_annotation(title = 'Winter Species Richness of Mobiles and Sessile Species in the Presence/Absence of Owl Limpets',
  #                theme = theme(plot.title = element_text(size = 15)))
```
####New Script For All Seasons####

#Data Manipulation for Mobile Species Abundance
```{r, Data Manipulation for Mobile Species Abundance}
cnm_mobiles_sum_spec <- mobiles_master %>%
  group_by(Season) %>% 
  mutate(spec_sum = rowSums(across(c(M_lugubris, L_hartwegii, L_dentiens, L_plena, L_keenae, L_limatula, L_scabra, Nuttallina_sp, Mopalia_sp, P_crassipes, T_eiseni, T_funebralis, T_gallina, L_gigantea)), na.rm=TRUE)) %>% 
   pivot_longer(cols = M_lugubris:L_gigantea, #pivoted to make longer
               names_to = "Species", #all species name in this column
               values_to = "Count") %>% 
  mutate(log_spec_sum = log(spec_sum+1))
```

#2-Boxplot of Mobile Organism Abundance Per Plot in Presence/Absence of Central Limpet
```{r,2-Boxplot of Mobile Organism Abundance Per Plot in Presence/Absence of Central Limpet}
 mob_boxplot <- cnm_mobiles_sum_spec %>% 
  ggplot(aes(
    x = Season,
    y = log_spec_sum,
    fill = COL_pres))+
  geom_boxplot()+
  labs(title = "Mobile Invertebrate Abundance in the Presence/Absence of a Central Owl Limpet",
       x = "Season",
       y = "Log Mobile Abundance",
       fill = "Owl Limpet Present")+
  scale_fill_brewer(palette = "Accent")+
  theme_bw()+
  theme(axis.text.x = element_text(size = 15),
        axis.title.x = element_text(size = 18, face = "bold"),
       axis.title.y = element_text(size = 18, face = "bold"))
mob_boxplot  
```

#General Linear Models for Mobile Abundance
R got stuck in this chunk, remove hashtags
```{r, General Linear Models for Mobile Abundance}
mob_abun_mod<-lmer(log_spec_sum~COL_pres*Season+(1|Transect), data = cnm_mobiles_sum_spec)
#plot(mob_abun_mod)
check_model(mob_abun_mod)
qqnorm(cnm_mobiles_sum_spec$log_spec_sum)
#anova(mob_abun_mod)
#summary(mob_abun_mod)

mob_abun_mod2<-glmer(spec_sum~COL_pres*Season+(1|Transect), family = "poisson"(link="log"), data = cnm_mobiles_sum_spec)
check_model(mob_abun_mod2)
check_overdispersion(mob_abun_mod2)
 abun_res <- simulateResiduals(mob_spec_rich_mod2)
    plot(abun_res) # Visual inspection for overdispersion
    testDispersion(abun_res) # Formal test
qqnorm(cnm_mobiles_sum_spec$spec_sum)
anova(mob_abun_mod2)
summary(mob_abun_mod2)
```

#Post Hoc test for Mobile Abundace
Remove hashtags, code works
```{r, Post Hoc test for Mobile Abundace}
emmeans(mob_abun_mod2, list(pairwise~COL_pres*Season), adjust = "tukey")
```

#Data Manipulation for Sessile Percent Cover Across Seasons
```{r, Sessile Species Abundance}
#Make columns into factors
#feb_cnm_sessile$COL_pres <- as.factor(feb_cnm_sessile$COL_pres) #made this column a factor THIS IMPORTANT FOR nMDS
#feb_cnm_sessile$Transect<- as.factor(feb_cnm_sessile$Transect)


cnm_sessile_sum_plot <- sess_percentcover %>%
  group_by(Season) %>% 
  mutate(spec_sum = rowSums(across(c(A_elegantissima, Phragmatopoma_californica, Chtalmus_Balanus, Bare_rock, Ralfsia_sp, Coralline_crust, Endocaldia_sp, P_polymerus,Ulva_sp)), na.rm=TRUE))%>% 
  mutate(Sessile_Invertebrate = (A_elegantissima+ Phragmatopoma_californica+ Chtalmus_Balanus+ P_polymerus)/spec_sum,
         Algae = (Ralfsia_sp+Coralline_crust+Endocaldia_sp+Ulva_sp)/spec_sum*100)

```

#Boxplot of Sessile Organism Percent Cover in Presence/Absence of Central Limpet Per Season
```{r, Boxplot of Sessile Organism Percent Cover in Presence/Absence of Central Limpet Per Season}
cnm_sessile_sum_plot$Season <- as.factor(cnm_sessile_sum_plot$Season)

sess_boxplot <- cnm_sessile_sum_plot %>%
  mutate(Season = factor(Season, levels = c("Fall", "Winter", "Spring"))) %>% 
  pivot_longer(cols = c(Sessile_Invertebrate, Algae), names_to = "Percent_Type")%>%
  ggplot(aes(
    x = COL_pres,
    y = value,
    fill = COL_pres))+
  geom_boxplot(show.legend = TRUE)+
  labs(x = "Presence/Absence of Central Owl Limpet",
       y = "Percent Cover",
       fill = "Percent Sessile Type")+
  scale_fill_brewer(palette = "Pastel1")+
  scale_color_brewer(palette = "Dark2")+
  theme_bw()+
  facet_grid(Percent_Type~Season)
sess_boxplot
```

#Liner Models for Sessile Animal Percent Cover
Chunk works, remove hasthags
```{r}
#mixed effects model
sess_ani_perc_mod <- lmer(Sessile_Invertebrate~COL_pres*Season+(1|Transect), data = cnm_sessile_sum_plot)
#plot(sess_ani_perc_mod)
check_model(sess_ani_perc_mod)
qqnorm(cnm_sessile_sum_plot$Sessile_Invertebrate)
var(cnm_sessile_sum_plot$Sessile_Invertebrate)
mean(cnm_sessile_sum_plot$Sessile_Invertebrate)
#anova(sess_ani_perc_mod)
#summary(sess_ani_perc_mod)

#genelized linear effects model
#using glmer.nb() implicitly uses a negative binomial distribution and a log link
sess_ani_perc_mod2 <-  glmer.nb(Sessile_Invertebrate~COL_pres*Season+(1|Transect),
                            control = glmerControl(optimizer = "bobyqa"),
                            data = cnm_sessile_sum_plot)
check_model(sess_ani_perc_mod2)
summary(sess_ani_perc_mod2)

#model_fit <- glmmTMB(Sessile_Invertebrate~COL_pres*Season+(1|Transect), data = cnm_sessile_sum_plot,
                    #family = beta_family(link = "logit"))

```

#Post Hoc test for Sessile Animal Percent Cover
Chunk works, remove hashtags
```{r}
emmeans(sess_ani_perc_mod2, list(pairwise~COL_pres*Season), adjust = "tukey")

```

#Liner Models for Sessile Algae Percent Cover
```{r}
#lmer
algae_perc_mod <- lmer(Algae~COL_pres*Season+(1|Transect), data = cnm_sessile_sum_plot)
check_model(algae_perc_mod)
qqnorm(cnm_sessile_sum_plot$Algae)
anova(algae_perc_mod)
summary(algae_perc_mod)

#using glmer.nb() implicitly uses a negative binomial distribution and a log link
algae_perc_mod2 <- glmer.nb(Algae~COL_pres*Season+(1|Transect),
                        data = cnm_sessile_sum_plot,
                        control = glmerControl(optimizer = "bobyqa"))
check_model(algae_perc_mod2)
summary(algae_perc_mod2)

```

#Post Hoc test for Algae Percent Cover
```{r, Post Hoc test for Algae Percent Cover}
emmeans(algae_perc_mod2, list(pairwise~COL_pres*Season), adjust = "tukey")
```

#Data Manipulation for Mobiles Species Richness
```{r, Data Manipulation for Mobiles Species Richness}
mob_count_mat_sum <- mobiles_matrix_master %>% 
  mutate(spec_sum = rowSums(across(c(M_lugubris, L_hartwegii, L_dentiens, L_plena, L_keenae, L_limatula, L_scabra, Nuttallina_sp, Mopalia_sp, P_crassipes, T_eiseni, T_funebralis, T_gallina, L_gigantea)), na.rm=TRUE)) %>% 
  mutate(log_spec_sum = log(spec_sum+1))

ggqqplot(mob_count_mat_sum$log_spec_sum)
qqnorm(mob_count_mat_sum$log_spec_sum)
```

#Boxplot of Mobiles Species Richness Across Seasons
```{r, Boxplot of Mobiles Species Richness Across Seasons}
mobile_spec_rich_box <- mob_count_mat_sum %>% 
  ggplot(aes(
    x = Season,
    y = spec_sum,
    fill = COL_pres,
    color = COL_pres))+
  geom_boxplot(show.legend = TRUE)+ 
  scale_fill_brewer(palette = "PuBuGn")+
  labs(title = "Mobile Invertabrate Species Richness in Presence/Absence of Central Owl Limpet",
       x = "Season",
       y = "Log Sum of Mobile Species Richness",
       color = "Owl Limpet Present",
       fill = "Owl Limpet Present")+
  #scale_x_discrete(limits = c("Fall", "Spring", "Winter"))
  theme_bw()+
  theme(axis.text.x = element_text(size = 15),
        axis.title.x = element_text(size = 18, face = "bold"),
       axis.title.y = element_text(size = 15, face = "bold"))
mobile_spec_rich_box
```

#General Linear Models for Mobile Species Richness
```{r, General Linear Models for Mobile Species Richness}
#lmer
mob_spec_rich_mod<-lmer(log_spec_sum~COL_pres*Season+(1|Transect), data = mob_count_mat_sum)
plot(mob_spec_rich_mod)
check_model(mob_spec_rich_mod)
qqnorm(mob_count_mat_sum$log_spec_sum)
anova(mob_spec_rich_mod)
summary(mob_spec_rich_mod)

#glmer
mob_spec_rich_mod2<-glmer(spec_sum~COL_pres*Season+(1|Transect), 
                          family = poisson(link = "log"),
                          data = mob_count_mat_sum)
 check_overdispersion(mob_spec_rich_mod2)
 sim_res <- simulateResiduals(mob_spec_rich_mod2)
    plot(sim_res) # Visual inspection for overdispersion
    testDispersion(sim_res) # Formal test
check_model(mob_spec_rich_mod2)
qqnorm(mob_count_mat_sum$spec_sum)
summary(mob_spec_rich_mod2)
model.matrix(mob_spec_rich_mod2, data = mob_count_mat_sum)
vif(mob_spec_rich_mod2)

#glmer negative binomial
mob_spec_rich_mod3<-glmer.nb(spec_sum~COL_pres*Season+(1|Transect), 
                          data = mob_count_mat_sum)
check_model(mob_spec_rich_mod3)
summary(mob_spec_rich_mod3)

```

#Post Hoc test for Mobile Species Richness
```{r, Post Hoc test for Mobile Species Richness}
emmeans(mob_spec_rich_mod2, list(pairwise~COL_pres*Season), adjust = "tukey")

```

#Data Manipulation For Sessile Species Richness
```{r}
sessile_spec_rich_matrix <- sessile_matrix_master %>%
  mutate(spec_sum = rowSums(across(c(A_elegantissima,Phragmatopoma_californica, Chtalmus_Balanus, Ralfsia_sp, Coralline_crust, Endocaldia_sp, P_polymerus, Ulva_sp)), na.rm=TRUE))%>% 
  mutate(Sessile_Invertebrate = (A_elegantissima+ Phragmatopoma_californica+ Chtalmus_Balanus+ P_polymerus),
         Algae = (Ralfsia_sp+Coralline_crust+Endocaldia_sp+Ulva_sp)) %>%  mutate(log_spec_sum = log(spec_sum+1),
         log_algae = log(Algae+1),
         log_sess_invert = log(Sessile_Invertebrate+1))

ggqqplot(sessile_spec_rich_matrix$log_spec_sum)

```

#Boxplot for Sessile Species Richness
```{r}
cnm_sessile_sum_plot$Season <- as.factor(cnm_sessile_sum_plot$Season)

sessile_spec_rich_boxplot <- sessile_spec_rich_matrix %>% 
  mutate(Season = factor(Season, levels = c("Fall", "Winter", "Spring"))) %>% 
  pivot_longer(cols = c(Sessile_Invertebrate, Algae), names_to = "sessile_type")%>%
  ggplot(aes(
    x = COL_pres,
    y = value,
    fill = sessile_type))+
  geom_boxplot(show.legend = TRUE)+ 
  scale_fill_brewer(palette = "Pastel1")+
  scale_color_brewer(palette = "Set1")+
  labs(x = "Presence/Absence of Central Owl Limpet",
       y = "Sum of Species Richness",
       fill = "Percent Sessile Type") +
  theme_bw()+
  facet_grid(sessile_type~Season)
sessile_spec_rich_boxplot

```

#Generalized Linear Models for Sessile Animals Species Richness
```{r, Generalized Linear Models for Sessile Animals Species Richness}
#lmer
sess_ani_spec_rich_mod<-lmer(log_sess_invert~COL_pres*Season+(1|Transect), data = sessile_spec_rich_matrix)
check_model(sess_ani_spec_rich_mod)
qqnorm(sessile_spec_rich_matrix$log_sess_invert)
ggqqplot(sessile_spec_rich_matrix$log_sess_invert)

#glmer
sess_ani_spec_rich_mod2<-glmer(Sessile_Invertebrate~COL_pres*Season+(1|Transect), 
                          family = poisson(link = "log"),
                          data = sessile_spec_rich_matrix)
check_model(sess_ani_spec_rich_mod2)
summary(sess_ani_spec_rich_mod2)
```

# Generalized Linear Models for Sessile Algae Species Richness
```{r, Generalized Linear Models for Sessile Algae Species Richness}
#lmer
sess_algae_spec_rich_mod<-lmer(log_algae~COL_pres*Season+(1|Transect), data = sessile_spec_rich_matrix)
check_model(sess_ani_spec_rich_mod)
qqnorm(sessile_spec_rich_matrix$log_sess_invert)
ggqqplot(sessile_spec_rich_matrix$log_sess_invert)
summary(sess_algae_spec_rich_mod)

#glmer with poisson distribution and log link
sess_algae_spec_rich_mod2<-glmer(Algae~COL_pres*Season+(1|Transect), 
                          family = poisson(link = "log"),
                          data = sessile_spec_rich_matrix)
check_model(sess_algae_spec_rich_mod2)
summary(sess_algae_spec_rich_mod2)

#glmer with negative binomial family
sess_algae_spec_rich_mod3 <- glmer.nb(Algae~COL_pres*Season+(1|Transect),
                        data = sessile_spec_rich_matrix,
                        control = glmerControl(optimizer = "bobyqa"))
summary(sess_algae_spec_rich_mod3)

```

# Post-Hoc Testing for Sessile Algae Species Richness
```{r, Post-Hoc Testing for Sessile Algae Species Richness}
emmeans(sess_algae_spec_rich_mod2, list(pairwise~COL_pres*Season), adjust = "tukey")

```

# Generalized Linear Model for All Sessile Species Richness
```{r, Generalized Linear Model for All Sessile Species Richness}
#lmer assuming normality
all_sess_spec_rich_mod<-lmer(log_spec_sum~COL_pres*Season+(1|Transect), data = sessile_spec_rich_matrix)
check_model(all_sess_spec_rich_mod)
summary(all_sess_spec_rich_mod)

#glmer for non-normal data
all_sess_spec_rich_mod2<-glmer(spec_sum~COL_pres*Season+(1|Transect), 
                          family = poisson(link = "log"),
                          data = sessile_spec_rich_matrix)
check_model(all_sess_spec_rich_mod2)
summary(all_sess_spec_rich_mod2)


```

# Post-Hoc Testing for All Sessile Species Richness
```{r, Post-Hoc Testing for All Sessile Species Richness}
emmeans(all_sess_spec_rich_mod2, list(pairwise~COL_pres*Season), adjust = "tukey")
```

#Patchwork for Mobile Species Abundance and Richness
```{r, Patchwork for Mobile Species Abundance and Richness}
patch_mob_abun <- cnm_mobiles_sum_spec %>% 
  ggplot(aes(
    x = Season,
    y = log_spec_sum,
    color = COL_pres,
    fill = COL_pres))+
  geom_boxplot()+
  scale_fill_brewer(palette = "Accent")+
  labs(x = "Season",
       y = "Log Mobile Species Abundance",
       fill = "Central OL Present",
       color = "Central OL Present") +
  theme_bw(base_size = 8)

patch_mob_spec_rich <- mob_count_mat_sum %>% 
  ggplot(aes(
    x = Season,
    y = log_spec_sum,
    fill = COL_pres,
    color = COL_pres))+
  geom_boxplot(show.legend = TRUE)+ 
  scale_fill_brewer(palette = "Dark2")+
  labs(x = "Season",
       y = "Log Mobile Species Richness",
       color = "Central OL Present",
       fill = "Central OL Present")+
  theme_bw(base_size = 8)

patch_mob_abun/patch_mob_spec_rich+
 plot_annotation(tag_levels = 'A')
```

